---
title: "Diversity_Exploration"
author: "Rachael Blake"
date: "October 29, 2015"
output: pdf_document
---

This is an initial exploration of diversity patterns in the shallow Groundfish areas in the central GOA.  This uses Ole's model output.


```{r,  include=FALSE, echo=FALSE, results='hide'}
library(httr)
library(plyr)
library(dplyr)
library(XML)
library(curl)
library(rvest)
library(tidyr)
library(stringr)
library(vegan)
library(reshape2)
library(ggplot2)
library(scales)
```


```{r,  include=FALSE, echo=FALSE, results='hide'}
URL_SpByArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uYmtIeFRIZzRtdlU"
SpByArea_Get <- GET(URL_SpByArea)
SpByArea_1 <- content(SpByArea_Get, as='text')
SpByArea <- read.csv(file=textConnection(SpByArea_1),stringsAsFactors=FALSE,head=TRUE)

```


```{r,  include=FALSE, echo=FALSE, results='hide'}
URL_SPCPUEArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uTkdHajFvbVVHc2c"
SPCPUEArea_Get <- GET(URL_SPCPUEArea)
SPCPUEArea_1 <- content(SPCPUEArea_Get, as='text')
SPCPUEArea <- read.csv(file=textConnection(SPCPUEArea_1),stringsAsFactors=FALSE,head=TRUE)

SPCPUEArea <- rename(SPCPUEArea, Year=year)

```


```{r, include=FALSE, echo=FALSE, results='hide'}
SpByArea <- SpByArea %>%
            mutate(Year = ifelse((year.numb=="1"),'1984',
                          ifelse((year.numb=="2"),'1987',
                          ifelse((year.numb=="3"),'1990',
                          ifelse((year.numb=="4"),'1993',       
                          ifelse((year.numb=="5"),'1996',
                          ifelse((year.numb=="6"),'1999',
                          ifelse((year.numb=="7"),'2001',
                          ifelse((year.numb=="8"),'2003',
                          ifelse((year.numb=="9"),'2005',
                          ifelse((year.numb=="10"),'2007',
                          ifelse((year.numb=="11"),'2009',
                          ifelse((year.numb=="12"),'2011',
                          ifelse((year.numb=="13"),'2013',
                          ifelse((year.numb=="14"),'2015',        
                          ""))))))))))))))
                   )
head(SpByArea)
```


-------------------------------------------------------
We want to look at diversity by area by year (using CPUE per area).
Here we calculate Shannon diversity, effective # of species, and beta diversity.

```{r, include=FALSE, echo=FALSE, results='hide'}
Div_SpCPUEArea <- SPCPUEArea %>%
                  filter(area!="Total") %>%
                  group_by(area, Year) %>% 
                  mutate(SW_Div = diversity(Mean.totalDensity, index="shannon"),
                         Exp_A_Div = exp(SW_Div)) %>%  # this is alpha diversity
                  ungroup() %>%
                  group_by(Year) %>%
                  mutate(G_Div = diversity(Mean.totalDensity, index="shannon"),
                         Exp_G_Div = exp(G_Div),
                         Exp_B_Div = (Exp_G_Div/Exp_A_Div)) %>%   # calculate beta diversity (gamma = alpha * beta)
                  ungroup() 

```
  
    
      
        
          
          
Plot of Shannon-Wiener Diversity:  

```{r, echo=FALSE, fig.height=4, fig.width=7}
SW <- ggplot(Div_SpCPUEArea, aes(Year, SW_Div, color=area)) + geom_line(size=1) + 
      scale_color_hue(breaks=c("1","2","3","4","5","6","7","8","9","10","11")) + 
      scale_x_continuous(breaks=pretty_breaks(n=15)) + 
      theme(panel.grid.minor=element_blank())
SW

```

\pagebreak
Plot of Alpha Diversity "exp(Shannon-Wiener)" ie: effective # of species :   

```{r, echo=FALSE, fig.height=4, fig.width=7}
Eff_Sp <- ggplot(Div_SpCPUEArea, aes(Year, Exp_A_Div, color=area)) +
          geom_line(size=1) + 
          scale_color_hue(breaks=c("1","2","3","4","5","6","7","8","9","10","11")) + 
          scale_x_continuous(breaks=pretty_breaks(n=15)) + 
          theme(panel.grid.minor=element_blank())
Eff_Sp

```

Plot beta diversity "(Exp_G_Div/Exp_A_Div)":   

```{r, echo=FALSE, fig.height=4, fig.width=7}
Bdiv <- ggplot(Div_SpCPUEArea, aes(Year, Exp_B_Div, color=area)) + geom_line(size=1) + 
        scale_color_hue(breaks=c("1","2","3","4","5","6","7","8","9","10","11")) + 
        scale_x_continuous(breaks=pretty_breaks(n=15)) + 
        theme(panel.grid.minor=element_blank())
Bdiv

```

\pagebreak
Plot gamma diversity:  

```{r, echo=FALSE, fig.height=4, fig.width=7}
Gdiv <- ggplot(Div_SpCPUEArea, aes(Year, Exp_G_Div)) + geom_line(size=1, color="blue") + 
        scale_color_hue(breaks=c("1","2","3","4","5","6","7","8","9","10","11")) + 
        scale_x_continuous(breaks=pretty_breaks(n=15)) + 
        theme(panel.grid.minor=element_blank())
Gdiv

```


```{r, eval=FALSE,include=FALSE}
mDiv_SpCPUEArea <- melt(Div_SpCPUEArea, id.vars = c("area", "Year"))
ggplot(mDiv_SpCPUEArea, aes(Year, value, group=variable))+geom_line()+facet_wrap(~area,ncol=2)

```

