---
title: "Diversity Paper Preliminary Plots"
author: "Mary Hunsicker"
date: "June 7, 2016"
output:
  pdf_document: 
  fig_height: 4
  fig_width: 7
  fig_caption: true
---

```{r, include=FALSE, echo=FALSE, results='hide'}
# Load all necessary packages
library(plyr) ; library(dplyr); library(tidyr) ; library(stringr) ; 
library(sp) ; library(maptools) ; library(raster) ; library(ggplot2) ; 
library(maps) ; library(mapdata) ; library(mapproj) ; library(httr) ; library(extrafont) ; 
library(vegan) ; library(scales)
```

```{r loadMapData, message=FALSE, echo=FALSE, warning=FALSE}
#Map of area locations
discrete_ShallowAreas<-read.csv("./goaTrawl/Output Data/Shallow_goa_discrete_areas_for_comparison_Combined(50_to_150m).MH.csv")
discrete_DeepAreas<-read.csv("./goaTrawl/Output Data/Deep_goa_discrete_areas_for_comparison_150_to_300m.MH.csv")
```

```{r map, cache=FALSE, echo=FALSE, warning=FALSE}
    ak<-map_data('worldHires','USA:Alaska')
    akmap<-ggplot()+geom_polygon(data=ak,aes(long,lat,group=group),fill=8,color="black")+
    theme(panel.background=element_rect(fill='aliceblue'))+
    xlab(expression(paste(Longitude^o,~'W')))+
    ylab(expression(paste(Latitude^o,~'N')))+
    coord_map(xlim = c(-160, -143),ylim = c(54, 61))
```

```{r, fig.cap="Locations of shallow areas (50-150 m) in Gulf of Alaska.",echo=FALSE,fig.height=3, fig.width=7}
#map of Shallow locations
discrete_ShallowAreas$area <- discrete_ShallowAreas$Area
discrete_ShallowAreas$Area <- factor(discrete_ShallowAreas$area)
akmap+geom_point(data=discrete_ShallowAreas,aes(LONGITUDE,LATITUDE,color=Area))
```

```{r, fig.cap="Locations of deep areas (>150-300 m) in Gulf of Alaska.",echo=FALSE,fig.height=3, fig.width=7}
#map of Deep locations
discrete_DeepAreas$area <- discrete_DeepAreas$Area
discrete_DeepAreas$Area <- factor(discrete_DeepAreas$area)
akmap+geom_point(data=discrete_DeepAreas,aes(LONGITUDE,LATITUDE,color=Area))
```

```{r,  include=FALSE, echo=FALSE, results='hide'}
# Load the Species by Area Occurrance data and clean it
URL_SpByArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uOGwySlN0T3NUQUE"
SpByArea_Get <- GET(URL_SpByArea)
SpByArea_1 <- content(SpByArea_Get, as='text')
SpByArea <- read.csv(file=textConnection(SpByArea_1), stringsAsFactors=FALSE, header=TRUE)


meanOcc <- SpByArea %>%
  mutate(Year = ifelse((year.numb=="1"),'1984',
                       ifelse((year.numb=="2"),'1987',
                              ifelse((year.numb=="3"),'1990',
                                     ifelse((year.numb=="4"),'1993',       
                                            ifelse((year.numb=="5"),'1996',
                                                   ifelse((year.numb=="6"),'1999',
                                                          ifelse((year.numb=="7"),'2001',
                                                                 ifelse((year.numb=="8"),'2003',
                                                                        ifelse((year.numb=="9"),'2005',
                                                                               ifelse((year.numb=="10"),'2007',
                                                                                      ifelse((year.numb=="11"),'2009',
                                                                                             ifelse((year.numb=="12"),'2011',
                                                                                                    ifelse((year.numb=="13"),'2013',
                                                                                                           ifelse((year.numb=="14"),'2015',        
                                                                                                                  ""))))))))))))))
  )
head(meanOcc)

```

```{r,  include=FALSE, echo=FALSE, results='hide'}
# Load the CPUE by Area data and merge in trawl species list
URL_SPCPUEArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uYzBOUFRtZklmX0U"
SPCPUEArea_Get <- GET(URL_SPCPUEArea)
SPCPUEArea_1 <- content(SPCPUEArea_Get, as='text')
SPCPUEArea <- read.csv(file=textConnection(SPCPUEArea_1), stringsAsFactors=FALSE, header=TRUE)

meanCPUE <- rename(SPCPUEArea, Year=year, Area=area)


```

```{r, include=FALSE, echo=FALSE, results='hide'}
# Read in deeper areas data CPUE
URL_deepCPUE <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uVF9VWnNPX3Z3S3c"
deepCPUE_Get <- GET(URL_deepCPUE)
deepCPUE_1 <- content(deepCPUE_Get, as='text')
deepCPUE <- read.csv(file=textConnection(deepCPUE_1),stringsAsFactors=FALSE,head=TRUE)

deepCPUE <- rename(deepCPUE, Year=year, Area=area)

```

```{r, include=FALSE, echo=FALSE, results='hide'}
# Read in deeper areas data OCCURRENCE
URL_deepOCC <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uZG1LbkphNmFpQ28"
deepOCC_Get <- GET(URL_deepOCC)
deepOCC_1 <- content(deepOCC_Get, as='text')
deepOCC_2 <- read.csv(file=textConnection(deepOCC_1),stringsAsFactors=FALSE,head=TRUE)


deepOCC <- deepOCC_2 %>%
  mutate(Year = ifelse((year.numb=="1"),'1984',
                       ifelse((year.numb=="2"),'1987',
                              ifelse((year.numb=="3"),'1990',
                                     ifelse((year.numb=="4"),'1993',       
                                            ifelse((year.numb=="5"),'1996',
                                                   ifelse((year.numb=="6"),'1999',
                                                          ifelse((year.numb=="7"),'2001',
                                                                 ifelse((year.numb=="8"),'2003',
                                                                        ifelse((year.numb=="9"),'2005',
                                                                               ifelse((year.numb=="10"),'2007',
                                                                                      ifelse((year.numb=="11"),'2009',
                                                                                             ifelse((year.numb=="12"),'2011',
                                                                                                    ifelse((year.numb=="13"),'2013',
                                                                                                           ifelse((year.numb=="14"),'2015',        
                                                                                                                  ""))))))))))))))
  )
head(deepOCC)


```

```{r, echo=FALSE}
# Make custom theme for plotting
theme_boxplot <- function(base_size = 12){
  theme_bw(base_size)%+replace%
    theme(legend.key.size=unit(12,"points"),
          legend.text=element_text(size=11),
          legend.key=element_blank(),
          legend.title=element_blank(),
          legend.background=element_rect(colour="white", fill="transparent"),
          plot.margin=unit(c(0.5,1,0.5,1), "lines"),  
          panel.border=element_blank(),
          panel.margin=unit(0,"lines"),
          panel.background=element_rect(fill=NA, colour=NA),
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
          axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'),
          axis.ticks.length=unit(1,"mm"),
          axis.text.x = element_text(margin=margin(5,0,0,0), size=12), 
          axis.text.y = element_text(margin=margin(0,5,0,0), size=12),
          axis.title.x=element_text(size=12, margin=margin(15,0,0,0)),
          axis.title.y=element_text(size=12, angle=90, margin=margin(0,15,0,0)),
          strip.text.x=element_text(size=12),
          strip.background=element_rect(colour="black", fill='white'))
}



```

```{r, include=FALSE, echo=FALSE}
# Functions to create plots
weight.var <- function(w){
  return(sum(w)^(-1))
}

space_plot <- function(df,NAME){
  par(mfrow=c(3,2),mar=c(3,4,1,0.5))
  AT <- df$Area  
  
  plot(1:5,1:5,type="n",axes=F,xlab="",ylab="")
  #text(1.5,4,NAME,cex=3,pos=4)
  y.lim <- c(min(df$Trend - df$SE.trend),max(df$Trend + df$SE.trend))
  plot(Trend~Area,data=df,ylim=y.lim,axes=F,pch=21,bg=1)
  abline(h=0,lty=2)
  arrows(x0=df$Area,x1=df$Area,
         y0=df$Trend + df$SE,y1= df$Trend - df$SE,
         length=0,lwd=1.5)
  axis(1,at=AT); box(bty="o")
  axis(2,las=2)
  title(xlab="Area",line=2)
  
  # B <- plot(Grand.w.mean~Area,data=df,type="b",axes=F,xlab="",pch=21,bg=1)
  #    axis(1,at=AT); box(bty="o")
  #    axis(2,las=2)
  #    title(xlab="Area",line=2)
  
  plot(SD~Area,data=df,type="b",axes=F,
       ylim= c(0,max(df$SD)),xlab="",pch=21,bg=1)
  axis(1,at=AT); box(bty="o")
  axis(2,las=2)
  title(xlab="Area",line=2)
  
  plot(CV~Area,data=df,type="b",axes=F,xlab="",pch=21,bg=1)
  axis(1,at=AT); box(bty="o")
  axis(2,las=2)
  title(xlab="Area",line=2)
  
  # E <- plot(Var~Grand.w.mean,data=df,axes=F,xlab="",pch=21,bg=1)
  #    axis(1); box(bty="o")
  #    axis(2,las=2)
  #    title(xlab="W.mean",line=2)
  
  # G <- c(A,C,D)
  
  return()
  
}



time_plot <- function(df,NAME){
  AREA <- as.numeric(as.character(unique(df$Area)))
  YEAR <- sort(unique(df$Year))
  y.lim <- c(min(df$Mean.totalDensity),max(df$Mean.totalDensity))
  
  for(i in 1:length(AREA)){
    if(AREA[i] !=1){par(new=T)}
    if(AREA[i] >= 3 & AREA[i] <= 5 ){
      plot(Mean.totalDensity~Year,data=df[df$Area ==AREA[i],],
           type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
    }
    
    if(AREA[i] < 3 | AREA[i] > 5){
      plot(Mean.totalDensity~Year,data=df[df$Area ==AREA[i],],
           type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
    }
  }
  
  axis(1,at=YEAR,las=1,hadj=0.6,tcl=-0.25)
  axis(2,las=2,hadj=0.5,tcl=-0.25)
  box(bty="l",lwd=2)
  #title(ylab=expression("Biomass (kg ha"^"-1"*")"),line=2.5)
  mtext(NAME,adj=0.05,line=-1)
  #abline(v=1989,lty=2,lwd=2)
  arrows(x0=1989,x1=1989,y0=y.lim[1],y1=y.lim[1]+
           c(y.lim[2]-y.lim[1])*0.8,length=0,lty=2,lwd=2)
}

##########
# Calculate least squares trend in each area and portfolio metrics.

trend_calc <- function(df,y,x, plot_name){
  vals <- NULL
  
  for(j in 1:length(AREA)){
    A <- lm(y~x,data=df[df$Area == AREA[j] & df$Year > 1989,]) 
    B <- lm(y~1,data=df[df$Area == AREA[j] & df$Year > 1989,]) 
    vals <- rbind(vals,c(AREA[j],
                         coef(summary(A))["x",c("Estimate","Std. Error")],
                         coef(summary(B))[1,c("Estimate","Std. Error")],
                         var(A$residuals),
                         sd(A$residuals)))
  }
  
  vals <- data.frame(vals)
  colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean",
                      "Grand.w.mean.se","Var","SD")
  vals$CV  <- vals$SD / vals$Grand.w.mean 
  
  df_2 <- vals
  
  C <- space_plot(df=df_2,NAME=plot_name)
  
  return(C)
}



#trend_calc(df=dat,y="SW_Div",x="Year", plot_name="Shannon Wiener")

#########

Plot_area_time <-function(df,y,ylabel){
  
  plot_1 <- ggplot(df, aes(x=Year, y=y, color=Area)) +
    geom_line(size=1) + theme_bw() + 
    ylab(ylabel) +
    geom_vline(aes(xintercept=1989), colour="black",
               linetype="dashed") +
    scale_color_manual(values=line_colors, 
                       breaks=area_breaks) + 
    scale_x_continuous(breaks=pretty_breaks(n=15)) + 
    theme(panel.grid=element_blank(), 
          legend.key=element_blank())
  
  return(plot_1)
} 


```

```{r, echo=FALSE, include=FALSE}
## Data for 9 deep areas
Div.Metrics<-read.csv("DiversityMetrics.Shallow.Bootstrapped2.csv")
Div.Metrics<-Div.Metrics[Div.Metrics$AREA!=10,]
Sp.Richness<-read.csv("Sp.Richness.Shallow.Bootstrapped.csv")
Sp.Richness<-Sp.Richness[Sp.Richness$AREA!=10,]
shallowDat<-cbind(Div.Metrics,Sp.Richness)
shallowDat<-shallowDat[,c(2:3,7:10,14)]

## Data for 5 deep areas
Div.Metrics<-read.csv("DiversityMetrics.Deep.Bootstrapped2.csv")
Div.Metrics<-Div.Metrics[Div.Metrics$AREA!=6,]
Sp.Richness<-read.csv("Sp.Richness.Deep.Bootstrapped.csv")
Sp.Richness<-Sp.Richness[Sp.Richness$AREA!=6,]
deepDat<-cbind(Div.Metrics,Sp.Richness)
deepDat<-deepDat[,c(2:3,7:10,14)]

shallowDat$AREA<-as.factor(shallowDat$AREA)
deepDat$AREA<-as.factor(deepDat$AREA)

##Boxplots Alpha Diversity Shallow Areas (exp H'):
sp_box1 <- ggplot(data=shallowDat, aes(AREA, y=Eff_Num_Sp)) + 
  geom_boxplot() + theme_boxplot() + xlab("Area (West <-> East)") +
  ylab("Alpha Diversity (exp H') - Shallow areas") +
  xlim("9","8","7","6","5","4","3","2","1") + 
  #  scale_fill_manual(values=barcolor) + ylim(0,0.5) +
  theme(legend.position="none", plot.background=element_blank(),
        axis.text.x=element_text(size=15),
        plot.title=element_text(colour="black", size=12,
                                hjust=0.04, vjust=0.5))
sp_box1

##Boxplots Alpha Diversity Deep Areas (exp H'):
sp_box2 <- ggplot(data=deepDat, aes(AREA, y=Eff_Num_Sp)) + 
  geom_boxplot() + theme_boxplot() + xlab("Area (West <-> East)") +
  ylab("Alpha Diversity (exp H') - Deep Areas") +
  xlim("5","4","3","2","1") +
  #  scale_fill_manual(values=barcolor) + ylim(0,0.5) +
  theme(legend.position="none", plot.background=element_blank(),
        axis.text.x=element_text(size=15),
        plot.title=element_text(colour="black", size=12,
                                hjust=0.04, vjust=0.5))
sp_box2

##Boxplots Beta Diversity Shallow Areas (exp H'):
sp_box3 <- ggplot(data=shallowDat, aes(AREA, y=Exp_B_Div)) + 
  geom_boxplot() + theme_boxplot() + xlab("Area (West <-> East)") +
  ylab("Beta Diversity - Shallow areas") +
  xlim("9","8","7","6","5","4","3","2","1") + 
  #  scale_fill_manual(values=barcolor) + ylim(0,0.5) +
  theme(legend.position="none", plot.background=element_blank(),
        axis.text.x=element_text(size=15),
        plot.title=element_text(colour="black", size=12,
                                hjust=0.04, vjust=0.5))
sp_box3

##Boxplots Beta Diversity Deep Areas (exp H'):
sp_box4 <- ggplot(data=deepDat, aes(AREA, y=Exp_B_Div)) + 
  geom_boxplot() + theme_boxplot() + xlab("Area (West <-> East)") +
  ylab("Alpha Diversity (exp H') - Deep Areas") +
  xlim("5","4","3","2","1") +
  #  scale_fill_manual(values=barcolor) + ylim(0,0.5) +
  theme(legend.position="none", plot.background=element_blank(),
        axis.text.x=element_text(size=15),
        plot.title=element_text(colour="black", size=12,
                                hjust=0.04, vjust=0.5))
sp_box4


##Boxplots Species Richness shallow areas:
sp_box5 <- ggplot(data=shallowDat, aes(AREA, y=Sp_rich)) + 
  geom_boxplot() + theme_boxplot() + xlab("Area (West <-> East)") +
  ylab("Species Richness - Shallow Areas") +
  xlim("9","8","7","6","5","4","3","2","1") + 
  #  scale_fill_manual(values=barcolor) + ylim(0,0.5) +
  theme(legend.position="none", plot.background=element_blank(),
        axis.text.x=element_text(size=15),
        plot.title=element_text(colour="black", size=12,
                                hjust=0.04, vjust=0.5))
sp_box5


##Boxplots Species Richness Deep areas:
sp_box6 <- ggplot(data=deepDat, aes(AREA, y=Sp_rich)) + 
  geom_boxplot() + theme_boxplot() + xlab("Area (West <-> East)") +
  ylab("Species Richness - Deep Areas") +
  xlim("5","4","3","2","1") + 
  #  scale_fill_manual(values=barcolor) + ylim(0,0.5) +
  theme(legend.position="none", plot.background=element_blank(),
        axis.text.x=element_text(size=15),
        plot.title=element_text(colour="black", size=12,
                                hjust=0.04, vjust=0.5))
sp_box6


##Boxplots Inverse Simpson shallow areas:
sp_box7 <- ggplot(data=shallowDat, aes(AREA, y=Invsimp)) + 
  geom_boxplot() + theme_boxplot() + xlab("Area (West <-> East)") +
  ylab("Inverse Simpson - Shallow Areas") +
  xlim("9","8","7","6","5","4","3","2","1") + 
  #  scale_fill_manual(values=barcolor) + ylim(0,0.5) +
  theme(legend.position="none", plot.background=element_blank(),
        axis.text.x=element_text(size=15),
        plot.title=element_text(colour="black", size=12,
                                hjust=0.04, vjust=0.5))
sp_box7


##Boxplots Inverse Simpson Deep areas:
sp_box8 <- ggplot(data=deepDat, aes(AREA, y=Invsimp)) + 
  geom_boxplot() + theme_boxplot() + xlab("Area (West <-> East)")+
  ylab("Inverse Simpson - Deep Areas")+
  xlim("5","4","3","2","1")+ 
  #  scale_fill_manual(values=barcolor) + ylim(0,0.5) +
  theme(legend.position="none", plot.background=element_blank(),
        axis.text.x=element_text(size=15),
        plot.title=element_text(colour="black", size=12,
                                hjust=0.04, vjust=0.5))
sp_box8
```


```{r, echo=FALSE, fig.height=4, fig.width=7,fig.cap="Alpha diversity (exp H') estimated for shallow areas (50-150m). Based on boostrap sampling (n=1000)."}
#{r sp_box1,fig.width=4,fig.height=3,echo=FALSE,message=FALSE,include=TRUE,fig.cap="Diversity#"}

sp_box1
```

```{r, echo=FALSE, fig.height=4, fig.width=7,fig.cap="Alpha Diversity (exp H') estimated for deep areas (>150-300m). Based on boostrap sampling (n=1000)."}
#{r sp_box1,fig.width=4,fig.height=3,echo=FALSE,message=FALSE,include=TRUE,fig.cap="Diversity#"}

sp_box2
```

```{r, echo=FALSE, fig.height=4, fig.width=7,fig.cap="Beta Diversity estimated for shallow areas (50-150m). Based on boostrap sampling (n=1000)."}
#{r sp_box1,fig.width=4,fig.height=3,echo=FALSE,message=FALSE,include=TRUE,fig.cap="Diversity#"}

sp_box3
```

```{r, echo=FALSE, fig.height=4, fig.width=7,fig.cap="Beta Diversity estimated for deep areas (>150-300m). Based on boostrap sampling (n=1000)."}
#{r sp_box1,fig.width=4,fig.height=3,echo=FALSE,message=FALSE,include=TRUE,fig.cap="Diversity#"}

sp_box4
```

```{r, echo=FALSE, fig.height=4, fig.width=7,fig.cap="Species richness estimated for shallow areas (50-150m). Based on boostrap sampling (n=1000)."}
#{r sp_box1,fig.width=4,fig.height=3,echo=FALSE,message=FALSE,include=TRUE,fig.cap="Diversity#"}

sp_box5
```

```{r, echo=FALSE, fig.height=4, fig.width=7,fig.cap="Species richness estimated for deep areas (>150-300m). Based on boostrap sampling (n=1000)."}
#{r sp_box1,fig.width=4,fig.height=3,echo=FALSE,message=FALSE,include=TRUE,fig.cap="Diversity#"}

sp_box6
```

```{r, echo=FALSE, fig.height=4, fig.width=7,fig.cap="Inverse Simpson's diversity estimated for shallow areas (50-150m). Based on boostrap sampling (n=1000)."}
#{r sp_box1,fig.width=4,fig.height=3,echo=FALSE,message=FALSE,include=TRUE,fig.cap="Diversity#"}

sp_box7
```

```{r fig6, echo=FALSE, fig.height=4, fig.width=7,fig.cap="Inverse Simpson's diversity estimated for deep areas (>150-300m). Based on boostrap sampling (n=1000)."}
#{r sp_box1,fig.width=4,fig.height=3,echo=FALSE,message=FALSE,include=TRUE,fig.cap="Diversity#"}

sp_box8
```

```{r, echo=FALSE}
# Calculate diversity metrics for shallower areas (50-150m)
Div_SpCPUEArea <- meanCPUE %>%
                  filter(Area!="Total") %>%
                  group_by(Area, Year) %>% 
                  mutate(SW_Div = diversity(Mean.totalDensity, index="shannon"),
                         Exp_A_Div = exp(SW_Div)) %>%  # this is alpha diversity
                  ungroup() %>%
                  group_by(Year) %>%
                  mutate(G_Div = diversity(Mean.totalDensity, index="shannon"),
                         Exp_G_Div = exp(G_Div),
                         Exp_B_Div = (Exp_G_Div/Exp_A_Div)) %>%   # calculate beta diversity (gamma = alpha * beta)
                  ungroup() %>%
                  group_by(Area, Year) %>%
                  mutate(Simp_Div = diversity(Mean.totalDensity, index="simpson"), 
                         InvSimp = diversity(Mean.totalDensity,index="inv")) %>%
                  ungroup %>%
                  arrange(Area, Year)
              
dat <- Div_SpCPUEArea
dat$Area <- as.factor(dat$Area)
datAREA <- sort(as.numeric(as.character(unique(dat$Area))))
datYEAR <- sort(unique(dat$Year))
```


```{r, echo=FALSE}
# Calculate diversity metrics for deeper areas (151-300m)
Div_SpCPUEDeep <- deepCPUE %>%
                  filter(Area!="Total") %>%
                  mutate_each(funs(as.character), Area) %>%
                  group_by(Area, Year) %>% 
                  mutate(SW_Div = diversity(Mean.totalDensity, index="shannon"),
                         Exp_A_Div = exp(SW_Div)) %>%  # this is alpha diversity
                  ungroup() %>%
                  group_by(Year) %>%
                  mutate(G_Div = diversity(Mean.totalDensity, index="shannon"),
                         Exp_G_Div = exp(G_Div),
                         Exp_B_Div = (Exp_G_Div/Exp_A_Div)) %>%   # calculate beta diversity (gamma = alpha * beta)
                  ungroup() %>%
                  group_by(Area, Year) %>%
                  mutate(Simp_Div = diversity(Mean.totalDensity, index="simpson"), 
                         InvSimp = diversity(Mean.totalDensity,index="inv")) %>%
                  ungroup %>%
                  arrange(Area, Year)

datDeep <- Div_SpCPUEDeep
datDeep$Area <- as.factor(datDeep$Area)
AREA <- sort(as.numeric(as.character(unique(datDeep$Area))))
YEAR <- sort(unique(datDeep$Year))
```

```{r, fig.cap="Time series of Alpha diversity for shallow areas (50-150m).", echo=FALSE, fig.height=4, fig.width=7}
#Plot Alpha Diversity Shallow Areas "exp(Shannon-Wiener)" ie: effective # of species
line_colors <- c("black","black","red","red","red","black","black","black","black")
area_breaks <- c("1","2","3","4","5","6","7","8","9")

Eff_Sp <- Plot_area_time(df=Div_SpCPUEArea, y=Div_SpCPUEArea$Exp_A_Div, 
                         ylabel="Alpha Diversity (exp H')-Shallow Areas")
Eff_Sp
```

```{r, fig.cap="Time series of Alpha diversity for deep areas (>150-300m).",echo=FALSE, fig.height=4, fig.width=7}
#Plot Alpha Diversity Deep Areas "exp(Shannon-Wiener)" ie: effective # of species
line_colors <- c("red","red","black","black","black")
area_breaks <- c("1","2","3","4","5")

Eff_Sp_dp <- Plot_area_time(df=Div_SpCPUEDeep, y=Div_SpCPUEDeep$Exp_A_Div, 
                            ylabel="Alpha Diversity (exp H')-Deep Areas")
Eff_Sp_dp
```

   
```{r, fig.cap="Time series of Beta diversity for shallow areas (50-150m).",echo=FALSE, fig.height=4, fig.width=7}
#Beta Diversity Shallow Areas "(Exp_G_Div/Exp_A_Div)":
line_colors <- c("black","black","red","red","red","black","black","black","black",
                 "black","black")
area_breaks <- c("1","2","3","4","5","6","7","8","9")

Bdiv <- Plot_area_time(df=Div_SpCPUEArea, y=Div_SpCPUEArea$Exp_B_Div,
                       ylabel="Beta Diversity-Shallow Areas")
Bdiv
```


```{r, fig.cap="Time series of Beta diversity for deep areas (>150-300m).",echo=FALSE, fig.height=4, fig.width=7}
#Beta Diversity Deep Areas "(Exp_G_Div/Exp_A_Div)": 
line_colors <- c("red","red","black","black","black")
area_breaks <- c("1","2","3","4","5")

Bdiv_dp <- Plot_area_time(df=Div_SpCPUEDeep, y=Div_SpCPUEDeep$Exp_B_Div,
                       ylabel="Beta Diversity-Deep Areas")
Bdiv_dp
```

```{r, echo=FALSE}
# Calculate Species Richness for Deep Areas: 
Div_SpOccArea <- meanOcc %>%
                 mutate_each(funs(as.numeric), Year) %>%
                 rename(Area=area) %>%
                 filter(Area!="Total") %>%
                 group_by(Area, Year) %>% 
                 mutate(SW_Div = diversity(Mean.avgPresence, index="shannon"),
                        Exp_A_Div = exp(SW_Div), # this is alpha diversity
                        Sp_Rich = sum(Mean.avgPresence)) %>%  
                 ungroup() %>%
                 group_by(Year) %>%
                 mutate(G_Div = diversity(Mean.avgPresence, index="shannon"),
                        Exp_G_Div = exp(G_Div),
                        Exp_B_Div = (Exp_G_Div/Exp_A_Div)) %>%   # calculate beta diversity (gamma = alpha * beta)
                 ungroup() 
              
dat_occ <-Div_SpOccArea
dat_occ$Year <- as.numeric(dat_occ$Year)
dat_occ$Area <- as.factor(dat_occ$Area)
datAREA <- sort(as.numeric(as.character(unique(dat_occ$Area))))
datYEAR <- sort(unique(dat_occ$Year))
```

```{r, echo=FALSE}
#Calculate Species Richness for Deep Areas:
Div_SpOccDeep <- deepOCC %>%
                 mutate_each(funs(as.numeric), Year) %>%
                 rename(Area=area) %>%
                 filter(Area!="Total") %>%
                 group_by(Area, Year) %>% 
                 mutate(SW_Div = diversity(Mean.avgPresence, index="shannon"),
                        Exp_A_Div = exp(SW_Div), 
                        Sp_Rich = sum(Mean.avgPresence)) %>%  # this is alpha diversity
                 ungroup() %>%
                 group_by(Year) %>%
                 mutate(G_Div = diversity(Mean.avgPresence, index="shannon"),
                        Exp_G_Div = exp(G_Div),
                        Exp_B_Div = (Exp_G_Div/Exp_A_Div)) %>%   # calculate beta diversity (gamma = alpha * beta)
                 ungroup() 
datDeep_occ <-Div_SpOccDeep
datDeep_occ$Year <- as.numeric(datDeep_occ$Year)
datDeep_occ$Area <- as.factor(datDeep_occ$Area)
AREA <- sort(as.numeric(as.character(unique(datDeep_occ$Area))))
YEAR <- sort(unique(datDeep_occ$Year))
```

```{r, fig.cap="Time series of Species Richness for shallow areas (50-150m).",echo=FALSE, fig.height=4, fig.width=7}
#Species Richness Shallow Areas:
# Shallow Species Richness
line_colors <- c("black","black","black","black","red","red","red","black","black")
area_breaks <- c("1","2","3","4","5","6","7","8","9")

SpRdiv <- Plot_area_time(df=Div_SpOccArea, y=Div_SpOccArea$Sp_Rich,
                         ylabel="Species Richness-Shallow Areas")
SpRdiv
```

```{r, fig.cap="Time series of Species Richness for deep areas (>150-300m).",echo=FALSE, fig.height=4, fig.width=7}
#Species richness Deep Areas: 
# Deep Species Richness
line_colors <- c("red","red","black","black","black")
area_breaks <- c("1","2","3","4","5")

SpRdeep <- Plot_area_time(df=Div_SpOccDeep, y=Div_SpOccDeep$Sp_Rich, 
                          ylabel="Species Richness-Deep Areas")
SpRdeep
```

\pagebreak

### References

Anderson, P. J., and J. F. Piatt. 1999. Community reorganization in the Gulf of Alaska following ocean climate        regime shift. Marine Ecology Progress Series 189:117-123.

Chesson, P., R. L. E. Gebauer, S. Schwinning, N. Huntly, K. Wiegand, M. S. K. Ernest, A. Sher, A. Novoplansky, j. F.   Weltzin. 2004. Resource pulses, species interactions, and diversity maintenance in arid and semi-arid               environments. Oecologia 141:236-253.

Gaichas, S., and R. C. Francis. 2008. Network models for ecosystem-based fishery analysis: a review of concepts and   application to the Gulf of alaska marine food web. Canadian Journal of Fisheries and Aquatic Science 65:1965-1982.

Gaichas, S. K., K. Y. Aydin, R. C. Francis, and J. Post. 2011. What drives dynamics in the Gulf of Alaska?            Integrating hypotheses of species, fishing, and climate relationships using ecosystem modeling. Canadian Journal    of Fisheries and Aquatic Sciences 68(9):1553-1578.

Hare, S. R., and N. J. Mantua. 2000. Empirical evidence for North Pacific regime shifts in 1977 and 1989. Progress    in Oceanography 47(2–4):103-145.

Hollowed, A. B., S. R. Hare, and Wooster, W. S. . 2001. Pacific basin climate variability and patterns of Northeast   Pacific marine fish production. Progress in Oceanography 49:257-282.

Hollowed, A. B., C. D. Wilson, P. J. Stabeno, and S. A. Salo. 2007. Effect of ocean conditions on the cross-shelf     distribution of walleye pollock (Theragra chalcogramma) and capelin (Malloutus villosus). Fisheries Oceanography    16(2):142-154.

Incardona, J. P., and coauthors. 2015. Very low embryonic crude oil exposures cause lasting cardiac defects in        salmon and herring. Scientific Reports 5:13499.

Jost, L. 2007. PARTITIONING DIVERSITY INTO INDEPENDENT ALPHA AND BETA COMPONENTS. Ecology 88(10):2427-2439.
  Lagerloef, G. 1983. Topographically controlled flow around a deep trough transecting the shelf off Kodiak Island,   Alaska. Journal of Physical Oceanography 13:139–146.

Logerwell, E. A., P. J. Stabeno, C. Wilson, and A. B. Hollowed. 2007. The effect of oceanographic variability and     interspecific competition on juvenile pollock and capelin distributions of the Gulf of Alaska Shelf. Deep Sea       Research II 54:2849-2686.

Ono, K., and coauthors. 2015. Space-time investigation of the effects of fishing on fish populations. Ecological      Applications.

Schindler, D. E., J. B. Armstrong, and T. E. Reed. 2015. The portfolio concept in ecology and evolution. Frontiers    in Ecology and the Environment 13(5):257-263.

Schindler, D. E., and coauthors. 2010. Population diversity and the portfolio effect in an exploited species. Nature   465(7298):609-612.

Shelton, A. O., J. T. Thorson, E. J. Ward, and B. E. Feist. 2014. Spatial semiparametric models improve estimates of   species abundance and distribution. Canadian Journal of Fisheries and Aquatic Sciences 71(11):1655-1666.

Smith, E. P., D. R. Orvos, and J. Cairns Jr. 1993. Impact Assessment Using the Before-After-Control-Impact (BACI)     Model: Concerns and Comments. Canadian Journal of Fisheries and Aquatic Sciences 50(3):627-637.

Stachura, M. M., and coauthors. 2014. Linking Northeast Pacific recruitment synchrony to environmental variability.   Fisheries Oceanography 23(5):389-408.

von Szalay, P. G., N. W. Raring, F. R. Shaw, M. E. Wilkins, and M. H. Martin. 2010 Data report: 2009 Gulf of Alaska   bottom trawl survey. , volume 208. U.S. Dep. Commer. , Seattle, WA.

Walline, P. D., C. D. Wilson, A. B. Hollowed, and S. C. Stienessen. 2012. Short-term effects of commercial fishing    on the distribution and abundance of walleye pollock (Theragra chalcogramma). Canadian Journal of Fisheries and     Aquatic Science 69:354-368.

Wilson, J. B. 2011. The twelve theories of co-existence in plant communities: the doubtful, the important and the     unexplored. Journal of Vegetation Science 22(1):184-195.


