---
title: "Groundfish nMDS"
author: "Colette Ward"
date: "May 5, 2016"
output: pdf_document
---

```{r, include=FALSE, echo=FALSE, results='hide'}
# Load packages
library(httr)
library(vegan)
library(plyr)
library(dplyr)
library(tidyr)

```

```{r,  include=FALSE, echo=FALSE, results='hide'}
# load & prepare Mean annual CPUE data for Shallow Areas:
URL_SPCPUEArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uYzBOUFRtZklmX0U" # new data for shallow areas
SPCPUEArea_Get <- GET(URL_SPCPUEArea)
SPCPUEArea_1 <- content(SPCPUEArea_Get, as='text')
SPCPUEArea <- read.csv(file=textConnection(SPCPUEArea_1),stringsAsFactors=FALSE,head=TRUE)
#View(SPCPUEArea)


cpue_spread <- SPCPUEArea %>%
  select(Species, Mean.totalDensity, area, year) %>%
  mutate(sp1 = substr(Species, 0,1), # extract the first letter of species name string
         sp2 = gsub(".*\\.","",Species), # remove everything before and including "."
         sp3 = paste(sp1, sp2, sep=".")) %>% # create shorter species names for ploting
  mutate(sp3 = gsub("D.Rockfish", "Sebastes1", sp3), 
         sp3 = gsub("R.Rockfish", "Sebastes2", sp3),
         sp3 = gsub("L.$", "Lepidopsetta", sp3),
         sp3 = gsub("M.Myctophidae$", "Myctophidae", sp3)) %>%
  select(-Species, -sp1, -sp2) %>% rename(Species = sp3) %>%
  spread(Species, Mean.totalDensity) %>% # make each species a column
  mutate(area = revalue(area, c("Total" = "10")), # recode Total for looping later
         area = as.numeric(area)) # convert to numeric class

shallowByArea_nmds_list <- split(cpue_spread, f = cpue_spread$area) # create a list of dataframes (one for each area; NB area 10 is Total)

shallowByArea_nmds_list1 <- lapply(shallowByArea_nmds_list, function(y){ row.names(y) <- y$year; y}) # create row names from year column
shallowByArea_nmds_list2 <- lapply(shallowByArea_nmds_list1, function(x) x[!(names(x) %in% c("area", "year"))]) # drop area & year

```

```{r,  include=FALSE, echo=FALSE, results='hide'}
# load & prepare mean annual CPUE for Deep areas:
URL_deepCPUE <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uVF9VWnNPX3Z3S3c"
deepCPUE_Get <- GET(URL_deepCPUE)
deepCPUE_1 <- content(deepCPUE_Get, as='text')
deepCPUE <- read.csv(file=textConnection(deepCPUE_1),stringsAsFactors=FALSE,head=TRUE)
#View(deepCPUE)

deepCPUE_spread <- deepCPUE %>%
  select(Species, Mean.totalDensity, area, year) %>%
  mutate(sp1 = substr(Species, 0,1),
         sp2 = gsub(".*\\.","",Species),
         sp3 = paste(sp1, sp2, sep=".")) %>%
  mutate(sp3 = gsub("D.Rockfish", "Sebastes1", sp3), 
         sp3 = gsub("R.Rockfish", "Sebastes2", sp3),
         sp3 = gsub("L.$", "Lepidopsetta", sp3),
         sp3 = gsub("M.Myctophidae$", "Myctophidae", sp3)) %>%
  select(-Species, -sp1, -sp2) %>% rename(Species = sp3) %>%
  spread(Species, Mean.totalDensity) %>% # make each species a column
  mutate(area = revalue(area, c("Total" = "6")), # recode Total for looping later
         area = as.numeric(area)) # convert to numeric class

deepByArea_nmds_list <- split(deepCPUE_spread, f = deepCPUE_spread$area) # create a list of dataframes (one for each area; NB area 6 is Total)

deepByArea_nmds_list1 <- lapply(deepByArea_nmds_list, function(y){ row.names(y) <- y$year; y}) # create row names from year column
deepByArea_nmds_list2 <- lapply(deepByArea_nmds_list1, function(x) x[!(names(x) %in% c("area", "year"))]) # drop area & year

```

######################################################
######################################################
######################################################

Transformation is Wisconsin-style double transformation (normalizes taxa to % abundance, then normalizes abundances to the maximum for each species) of sqrt-transformed data.

Sorry the text is so small in the ordination plots, need small text for legibility because there are so many species. 
Enlarge the output pdf to see species names.


```{r,  include=FALSE, echo=FALSE, results='hide'}

# nMDS for Shallow Areas:
ordShallow <- list()
for (i in seq_along(shallowByArea_nmds_list2)) {
  ordShallow[[i]] <- metaMDS(shallowByArea_nmds_list2[[i]], distance='bray', trymax=1000)
}


# examine Shepard's stressplots:
# goodness-of-fit test: plots ordination distances against community dissimilarities
par (mfrow = c(3,4), pty="m") 
stressplotsShallow <- list()
for (i in seq_along(ordShallow)) {
  stressplotsShallow[[i]] <- stressplot(ordShallow[[i]], main = paste("Area",i))
}


# examine goodness of fit plots:
# size represents goodness of fit (bigger = worse fit)
par (mfrow = c(3,4), pty="m") 
goodnessShallow <- list()
for (i in seq_along(ordShallow)) {
  goodnessShallow[[i]] <- plot (ordShallow[[i]], display = 'sites', type = 't', main = paste("Area",i))
  points (ordShallow[[i]], display = 'sites', cex = goodness (ordShallow[[i]])*200)
}

```

#Shallow areas
Area 10 is Total

```{r, echo=FALSE}
# Ordination plots:
par(mfrow = c(2,2), mar=c(2,2,2,2))
ordPlotsShallow1 <- list()
for (i in seq_along(ordShallow)) {
  ordPlotsShallow1[[i]] <- plot(ordShallow[[i]], type = "n", main = paste("Shallow Area",i))
  text(ordShallow[[i]], display = "spec", cex = 0.25, col="red")
  text(ordShallow[[i]], display = "sites", cex=0.7, col="blue")
}
```

***


```{r,  include=FALSE, echo=FALSE, results='hide'}

# nMDS for Deep Areas:
ordDeep <- list()
for (i in seq_along(deepByArea_nmds_list2)) {
  ordDeep[[i]] <- metaMDS(deepByArea_nmds_list2[[i]], distance='bray', trymax=1000)
}


# examine Shepard's stressplots:
# goodness-of-fit test: plots ordination distances against community dissimilarities
par (mfrow = c(2,3), pty="m") 
stressplotsDeep <- list()
for (i in seq_along(ordDeep)) {
  stressplotsDeep[[i]] <- stressplot(ordDeep[[i]], main = paste("Area",i))
}


# examine goodness of fit plots:
# size represents goodness of fit (bigger = worse fit)
par (mfrow = c(2,3), pty="m") 
goodnessDeep <- list()
for (i in seq_along(ordDeep)) {
  goodnessDeep[[i]] <- plot (ordDeep[[i]], display = 'sites', type = 't', main = paste("Area",i))
  points (ordDeep[[i]], display = 'sites', cex = goodness (ordDeep[[i]])*200)
}

```

#Deep areas
Area 6 is Total

```{r, echo=FALSE}
# Ordination plots:
# NB Area 6 is total of our 5 areas
par(mfrow = c(2,2), mar=c(2,2,2,2))
ordPlotsDeep1 <- list()
for (i in seq_along(ordDeep)) {
  ordPlotsDeep1[[i]] <- plot(ordDeep[[i]], type = "n", main = paste("Deep Area",i))
  text(ordDeep[[i]], display = "spec", cex = 0.25, col="red")
  text(ordDeep[[i]], display = "sites", cex=0.7, col="blue")
}

```
