---
title: "Groundfish multivariate analyses - Comparing Areas By Year"
author: "Colette Ward"
date: "May 6, 2016"
output: pdf_document
---

```{r, include=FALSE, echo=FALSE, results='hide'}
# Load packages
library(httr)
library(vegan)
library(mvnormtest)
library(plyr)
library(dplyr)
library(tidyr)

```


```{r,  include=FALSE, echo=FALSE, results='hide'}
# load & prep look-up table of common names
URL_common <- "https://drive.google.com/uc?export=download&id=0B1XbkXxdfD7ubzIzdURzanhfZnc" # this is "trawl_species_control_file.csv"
common_Get <- GET(URL_common)
common_1 <- content(common_Get, as='text')
common <- read.csv(file=textConnection(common_1),stringsAsFactors=FALSE,head=TRUE)

common1 <- common %>%
  select(database.name, common.name) %>%
  rename(Species = database.name)
for (i in 1:nrow(common1)) { # add common names for Sebastes 1 & 2
  if(common1$Species[i] == "Dusky.and.Dark.Rockfish") {common1$common.name[i] <- "Sebastes 1"}
  if(common1$Species[i] == "Rougheye.and.Blackspotted.Rockfish") {common1$common.name[i] <- "Sebastes 2"}
}
```


```{r,  include=FALSE, echo=FALSE, results='hide'}
# load mean annual CPUE data for Shallow Areas:
#URL_shallowCPUEArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uYzBOUFRtZklmX0U" # new data for shallow areas
#shallowCPUEArea_Get <- GET(URL_shallowCPUEArea)
#shallowCPUEArea_1 <- content(shallowCPUEArea_Get, as='text')
#shallowCPUEArea <- read.csv(file=textConnection(shallowCPUEArea_1),stringsAsFactors=FALSE,head=TRUE)
#View(shallowCPUEArea)

# load mean annual CPUE data for Shallow Areas (these are Ole's 11 areas)
setwd("~/Google Drive/GoA project/pfx-groundfish/diversity-data")
shallowCPUEArea <- read.csv("All_sp_index_meanCPUEByArea.csv", header = T, stringsAsFactors = FALSE)

shallowCPUEArea2 <- left_join(shallowCPUEArea, common1, by = "Species") %>% # merge common names onto SPCPUEArea
  filter(area != "Total", area != "8") %>% # remove regional totals and Ole's area 8
  mutate(area = revalue(area, c("9"="8", "10"="9", "11"="10"))) # renumber (old = new) shallow areas to account for splitting area 7 into 7, 8, 9 (but removing 8)

```

```{r,  include=FALSE, echo=FALSE, results='hide'}
# load mean annual CPUE for Deep areas:
URL_deepCPUEArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uVF9VWnNPX3Z3S3c"
deepCPUEArea_Get <- GET(URL_deepCPUEArea)
deepCPUEArea_1 <- content(deepCPUEArea_Get, as='text')
deepCPUEArea <- read.csv(file=textConnection(deepCPUEArea_1),stringsAsFactors=FALSE,head=TRUE)
#View(deepCPUEArea)

deepCPUEArea2 <- left_join(deepCPUEArea, common1, by = "Species") %>% # merge in common names
  filter(area != "Total") %>%
  mutate(area = revalue(area, c("1"="11", "2"="12", "3"="14", "4"="15", "5"="13"))) # renumber (old = new) shallow areas to account for splitting area 7 into 7, 8, 9 (but removing 8)

```

```{r, echo=FALSE}
#which are the most dominant species?
s <- shallowCPUEArea2 %>%
  filter(area == "Total") %>%
  group_by(common.name) %>%
  summarize(sum(Mean.totalDensity)) %>%
  ungroup() %>%
  arrange(desc(`sum(Mean.totalDensity)`))
print(head(s))
#1 arrowtooth flounder                 900.2215
#2     Pacific halibut                 338.6928
#3         Pacific cod                 268.7385
#4     walleye pollock                 247.9965
#5 Pacific ocean perch                 165.1791
#6   northern rockfish                 109.9819



d <- deepCPUEArea2 %>%
  filter(area == "Total") %>%
  group_by(common.name) %>%
  summarize(sum(Mean.totalDensity)) %>%
  ungroup() %>%
  arrange(desc(`sum(Mean.totalDensity)`))
print(head(d))
# 1 arrowtooth flounder                1367.5554
# 2 Pacific ocean perch                 645.4675
# 3     walleye pollock                 393.4510
# 4           sablefish                 204.2194
# 5       flathead sole                 122.4128
# 6         Pacific cod                 117.3530

```

######################################################
######################################################
######################################################

# nMDS
Transformation is Wisconsin-style double transformation (normalizes taxa to % abundance, then normalizes abundances to the maximum for each species) of sqrt-transformed data.

We do not use a Hellinger transformation because it reduces the influence of very rare species and inflates the influence of very abundant species (renders the result driven by Arrowtooth).


```{r,  include=FALSE, echo=FALSE, results='hide'}

# organize Shallow areas data for nMDS:
shallowCPUE_spread3 <- shallowCPUEArea2 %>%
  select(common.name, Mean.totalDensity, area, year) %>%
  spread(common.name, Mean.totalDensity) %>% # make each species a column
  mutate(Area = as.numeric(area)) %>% select(-area) # convert area to numeric



# organize Deep areas data for nMDS:
deepCPUE_spread3 <- deepCPUEArea2 %>%
  select(common.name, Mean.totalDensity, area, year) %>%
  spread(common.name, Mean.totalDensity) %>% # make each species a column
  mutate(Area = as.numeric(area)) %>% select(-area) # convert area to numeric



# alternative: do nMDS on Shallow & Deep areas combined:
CPUE_spread3 <- rbind(shallowCPUE_spread3, deepCPUE_spread3)



#####################

# run the nMDS:
# Shallow areas:

ordShallowWiscSqrt <- metaMDS(shallowCPUE_spread3[2:54], distance='bray', trymax=1000)
ordShallowHellinger <- metaMDS(decostand(shallowCPUE_spread3[,2:54], "hell"), distance='bray', trymax=1000) # goodness of fit plots look better on data with Hellinger transformation
# also considered method = "log" (does not converge) and "max" (results are similar to wisc(sqrt( )))

par(mfrow=c(1,1))
stressplot(ordShallowWiscSqrt) # make sure stressplots show good fits

plot(ordShallowWiscSqrt, display = 'sites', type = 't', main = "Shallow Areas") # look at goodness-of-fit plots
points(ordShallowWiscSqrt, display = 'sites', cex = goodness (ordShallowWiscSqrt)*200)

  
  
# Deep areas:
ordDeepWiscSqrt <- metaMDS(deepCPUE_spread3[2:54], distance='bray', trymax=1000)
ordDeepHellinger <- metaMDS(decostand(deepCPUE_spread3[2:54],"hell"), distance='bray', trymax=1000)

stressplot(ordDeepWiscSqrt) # make sure stressplots show good fits

plot(ordDeepWiscSqrt, display = 'sites', type = 't', main = "Deep Areas") # look at goodness-of-fit plots
points(ordDeepWiscSqrt, display = 'sites', cex = goodness (ordDeepWiscSqrt)*200)



# Shallow & Deep areas combined:
ordAllWiscSqrt <- metaMDS(CPUE_spread3[2:54], distance='bray', trymax=1000)
ordAllHellinger <- metaMDS(decostand(CPUE_spread3[2:54],"hell"), distance='bray', trymax=1000) # many runs (678) before a solution was reached

stressplot(ordAllWiscSqrt) # make sure stressplots show good fits
stressplot(ordAllHellinger) # make sure stressplots show good fits

plot(ordAllWiscSqrt, display = 'sites', type = 't') # look at goodness-of-fit plots
points(ordAllWiscSqrt, display = 'sites', cex = goodness (ordAllWiscSqrt)*200)

plot(ordAllHellinger, display = 'sites', type = 't') # look at goodness-of-fit plots
points(ordAllHellinger, display = 'sites', cex = goodness (ordAllHellinger)*200)

```


```{r, echo=FALSE}

# Ordination plots with grouping by areas:


# Create ordination plots for Wisconsin double standardization + Square root transformation:

# Plot shallow areas:
ordiplot(ordShallowWiscSqrt, type="n", choices=c(1,2), main=NULL, xlab="NMDS 1", ylab="NMDS 2")#, ylim = c(-0.3, 0.2))
#title(main="Shallow Areas, Wisconsin + Sqrt")
#text(ordShallowWiscSqrt, display = "sites", cex = 0.5, col="black")
text(ordShallowWiscSqrt, display = "spec", cex = 0.5, col="black")
#ordiplot(ordShallowWiscSqrt, display = "sites") #, groups=shallowCPUE_spread3$Area)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$Area, show.groups=1, col="red", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$Area, show.groups=2, col="orange", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$Area, show.groups=3, col="goldenrod", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$Area, show.groups=4, col="brown", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$Area, show.groups=5, col="green", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$Area, show.groups=6, col="blue", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$Area, show.groups=7, col="purple", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$Area, show.groups=8, col="pink", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$Area, show.groups=9, col="dark grey", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$Area, show.groups=10, col="black", label=T, lwd=3, lty=1)


# areas 8 & 9  are very different from the rest, and also from each other.
# there is also an east-west gradient in areas 1-7.
# and area 1 does not overlap with any others.
# is area 7 similar to other areas outside the Alaska Coastal Current?
# note that area 8 & 9 hulls look smaller than hulls for areas 1-4 & 7 ... suggests that temporal turnover in 8 & 9 was less than that in 1-4 & 7 ... use betadisper test (must use same transformation) and then anova(test) to test this. Only use betadisper on factors that permanova says are significant.


# Plot deep areas:
ordiplot(ordDeepWiscSqrt, type="n", choices=c(1,2), main=NULL, xlab="NMDS 1", ylab="NMDS 2")#, ylim = c(-0.3, 0.2))
#text(ordDeepWiscSqrt, display = "sites", cex = 0.5, col="black")
#title(main="Deep Areas, Wisconsin + Sqrt")
text(ordDeepWiscSqrt, display = "spec", cex = 0.5, col="black")
#ordiplot(ordDeepWiscSqrt, display = "sites") #, groups=deepCPUE_spread3$Area)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$Area, show.groups=11, col="red", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$Area, show.groups=12, col="orange", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$Area, show.groups=13, col="brown", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$Area, show.groups=14, col="green", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$Area, show.groups=15, col="blue", label=T, lwd=3, lty=1)




###############

# Ordination plot for shallow + deep areas, Wisconsin-Sqrt Transformation:
ordiplot(ordAllWiscSqrt, type="n", choices=c(1,2), main="Wisc-Sqrt Transformation", xlab="NMDS 1", ylab="NMDS 2")#, ylim = c(-0.3, 0.2))
#title(main="Wisconsin + Sqrt Transformation")
#text(ordAllWiscSqrt, display = "sites", cex = 0.5, col="black")
text(ordAllWiscSqrt, display = "spec", cex = 0.5, col="black")
#ordiplot(ordAllWiscSqrt, display = "sites") #, groups=CPUE_spread3$Area)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=1, col="red", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=2, col="orange", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=3, col="goldenrod", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=4, col="brown", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=5, col="green", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=6, col="blue", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=7, col="purple", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=8, col="pink", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=9, col="dark grey", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=10, col="black", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=11, col="springgreen4", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=12, col="steelblue3", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=13, col="magenta", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=14, col="tomato1", label=T, lwd=3, lty=1)
ordihull(ordAllWiscSqrt, groups=CPUE_spread3$Area, show.groups=15, col="orangered4", label=T, lwd=3, lty=1)




#####

# Create ordination plots for Hellinger transformation:

# Plot shallow areas:
ordiplot(ordShallowHellinger, type="n", choices=c(1,2), xlab="NMDS 1", ylab="NMDS 2")#, ylim = c(-0.3, 0.2)), main="Hellinger Transformation",
#title(main="Shallow Areas, Hellinger")
#text(ordShallowHellinger, display = "sites", cex = 0.5, col="black")
text(ordShallowHellinger, display = "spec", cex = 0.5, col="black")
#ordiplot(ordShallowHellinger, display = "sites") #, groups=shallowCPUE_spread3$Area)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$Area, show.groups=1, col="red", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$Area, show.groups=2, col="orange", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$Area, show.groups=3, col="goldenrod", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$Area, show.groups=4, col="brown", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$Area, show.groups=5, col="green", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$Area, show.groups=6, col="blue", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$Area, show.groups=7, col="purple", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$Area, show.groups=8, col="pink", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$Area, show.groups=9, col="dark grey", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$Area, show.groups=10, col="black", label=T, lwd=3, lty=1)



# Plot deep areas:
ordiplot(ordDeepHellinger, type="n", choices=c(1,2), main=NULL, xlab="NMDS 1", ylab="NMDS 2")#, ylim = c(-0.3, 0.2))
#text(ordShallowHellinger, display = "sites", cex = 0.5, col="black")
#title(main="Deep Areas, Hellinger")
text(ordDeepHellinger, display = "spec", cex = 0.5, col="black")
#ordiplot(ordDeepHellinger, display = "sites") #, groups=deepCPUE_spread3$Area)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$Area, show.groups=11, col="red", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$Area, show.groups=12, col="orange", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$Area, show.groups=13, col="brown", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$Area, show.groups=14, col="green", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$Area, show.groups=15, col="blue", label=T, lwd=3, lty=1)




#############

# Ordination plot for shallow + deep areas, Hellinger Transformation:
ordiplot(ordAllHellinger, type="n", choices=c(1,2), xlab="NMDS 1", ylab="NMDS 2")#, ylim = c(-0.3, 0.2))
#title(main="Hellinger Transformation")
#text(ordAllHellinger, display = "sites", cex = 0.5, col="black")
text(ordAllHellinger, display = "spec", cex = 0.5, col="black")
#ordiplot(ordAllHellinger, display = "sites") #, groups=CPUE_spread3$Area)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=1, col="red", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=2, col="orange", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=3, col="goldenrod", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=4, col="brown", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=5, col="green", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=6, col="blue", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=7, col="purple", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=8, col="pink", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=9, col="dark grey", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=10, col="black", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=11, col="springgreen4", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=12, col="steelblue3", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=13, col="magenta", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=14, col="tomato1", label=T, lwd=3, lty=1)
ordihull(ordAllHellinger, groups=CPUE_spread3$Area, show.groups=15, col="orangered4", label=T, lwd=3, lty=1)

```



``` {r}
# Ordination plots with grouping by years:

# Wisconsin transformation:
par(mfrow=c(1,2))
ordiplot(ordShallowWiscSqrt, type="n", choices=c(1,2), main=NULL, xlab="NMDS 1", ylab="NMDS 2")#, ylim = c(-0.3, 0.2))
title(main="Shallow Areas")
#text(ordShallowWiscSqrt, display = "sites", cex = 0.5, col="black")
text(ordShallowWiscSqrt, display = "spec", cex = 0.5, col="black")
#ordiplot(ordShallowWiscSqrt, display = "sites") #, groups=shallowCPUE_spread3$Area)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=1984, col="red", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=1987, col="orange", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=1990, col="yellow", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=1993, col="green", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=1996, col="dark green", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=1999, col="turquoise", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=2001, col="blue", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=2003, col="dark blue", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=2005, col="lavender", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=2007, col="purple", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=2009, col="pink", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=2011, col="light grey", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=2013, col="dark grey", label=T, lwd=3, lty=1)
ordihull(ordShallowWiscSqrt, groups=shallowCPUE_spread3$year, show.groups=2015, col="black", label=T, lwd=3, lty=1)




ordiplot(ordDeepWiscSqrt, type="n", choices=c(1,2), main=NULL, xlab="NMDS 1", ylab="NMDS 2")#, ylim = c(-0.3, 0.2))
#text((ordDeepWiscSqrt,, display = "sites", cex = 0.5, col="black")
title(main="Deep Areas")
text(ordDeepWiscSqrt, display = "spec", cex = 0.5, col="black")
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=1984, col="red", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=1987, col="orange", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=1990, col="yellow", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=1993, col="green", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=1996, col="dark green", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=1999, col="turquoise", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=2001, col="blue", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=2003, col="dark blue", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=2005, col="lavender", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=2007, col="purple", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=2009, col="pink", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=2011, col="light grey", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=2013, col="dark grey", label=T, lwd=3, lty=1)
ordihull(ordDeepWiscSqrt, groups=deepCPUE_spread3$year, show.groups=2015, col="black", label=T, lwd=3, lty=1)





# Hellinger Transformation:
par(mfrow=c(1,2))
ordiplot(ordShallowHellinger, type="n", choices=c(1,2), main=NULL, xlab="NMDS 1", ylab="NMDS 2")#, ylim = c(-0.3, 0.2))
title(main="Shallow Areas")
#text(ordShallowHellinger, display = "sites", cex = 0.5, col="black")
text(ordShallowHellinger, display = "spec", cex = 0.5, col="black")
#ordiplot(ordShallowHellinger, display = "sites") #, groups=shallowCPUE_spread3$Area)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=1984, col="red", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=1987, col="orange", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=1990, col="yellow", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=1993, col="green", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=1996, col="dark green", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=1999, col="turquoise", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=2001, col="blue", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=2003, col="dark blue", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=2005, col="lavender", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=2007, col="purple", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=2009, col="pink", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=2011, col="light grey", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=2013, col="dark grey", label=T, lwd=3, lty=1)
ordihull(ordShallowHellinger, groups=shallowCPUE_spread3$year, show.groups=2015, col="black", label=T, lwd=3, lty=1)




ordiplot(ordDeepHellinger, type="n", choices=c(1,2), main=NULL, xlab="NMDS 1", ylab="NMDS 2")#, ylim = c(-0.3, 0.2))
#text(ordDeepHellinger, display = "sites", cex = 0.5, col="black")
title(main="Deep Areas")
text(ordDeepHellinger, display = "spec", cex = 0.5, col="black")
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=1984, col="red", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=1987, col="orange", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=1990, col="yellow", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=1993, col="green", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=1996, col="dark green", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=1999, col="turquoise", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=2001, col="blue", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=2003, col="dark blue", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=2005, col="lavender", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=2007, col="purple", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=2009, col="pink", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=2011, col="light grey", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=2013, col="dark grey", label=T, lwd=3, lty=1)
ordihull(ordDeepHellinger, groups=deepCPUE_spread3$year, show.groups=2015, col="black", label=T, lwd=3, lty=1)

```

######################################################
######################################################
######################################################


# PERMANOVA

```{r,  include=FALSE, echo=FALSE}
# must still add oil-exposed vs unexposed groupings

# Shallow areas:
adonis(wisconsin(sqrt(shallowCPUE_spread3[,2:54])) ~ Area*year, data = shallowCPUE_spread3, permutations = 200, method = "bray")
#           Df SumsOfSqs MeanSqs F.Model      R2   Pr(>F)   
#Area        1    1.2116 1.21156  45.282 0.21624 0.004975 **
#year        1    0.6860 0.68597  25.638 0.12243 0.004975 **
#Area:year   1    0.0665 0.06655   2.487 0.01188 0.054726 . 
#Residuals 136    3.6388 0.02676         0.64945            
#Total     139    5.6029                 1.00000 


adonis(wisconsin(sqrt(shallowCPUE_spread3[,2:54])) ~ Area, data = shallowCPUE_spread3, permutations = 200, method = "bray")
#           Df SumsOfSqs MeanSqs F.Model      R2   Pr(>F)   
#Area        1    1.2116 1.21156  38.074 0.21624 0.004975 **
#Residuals 138    4.3914 0.03182         0.78376            
#Total     139    5.6029                 1.00000 


adonis(wisconsin(sqrt(shallowCPUE_spread3[,2:54])) ~ Area + year, data = shallowCPUE_spread3, permutations = 200, method = "jaccard")
# Terms added sequentially (first to last)
#           Df SumsOfSqs MeanSqs F.Model      R2   Pr(>F)   
#Area        1    2.3476 2.34762  34.120 0.17900 0.004975 **
#year        1    1.3410 1.34096  19.489 0.10225 0.004975 **
#Residuals 137    9.4263 0.06881         0.71875            
#Total     139   13.1149                 1.00000 


######################################################

# Using Hellinger transformation:
adonis(decostand(shallowCPUE_spread3[2:54],"hell") ~ Area*year, data = shallowCPUE_spread3, permutations = 200, method = "bray")
# Note this significance is likely driven by differences between 8 & 9 vs all other sites, not sites 1-6
#           Df SumsOfSqs MeanSqs F.Model      R2   Pr(>F)   
#Area        1    0.8614 0.86138  39.894 0.20597 0.004975 **
#year        1    0.3261 0.32614  15.105 0.07799 0.004975 **
#Area:year   1    0.0579 0.05794   2.683 0.01385 0.039801 * 
#Residuals 136    2.9365 0.02159         0.70218            
#Total     139    4.1820                 1.00000   


adonis(decostand(deepCPUE_spread3[2:54],"hell") ~ Area*year, data = deepCPUE_spread3, permutations = 200, method = "bray")
# Note this significance is likely driven by differences between 5 vs all other sites, not sites 1-4
#          Df SumsOfSqs MeanSqs F.Model      R2   Pr(>F)   
#Area       1   0.13534 0.13534  8.8843 0.08887 0.004975 **
#year       1   0.35357 0.35357 23.2101 0.23217 0.004975 **
#Area:year  1   0.02860 0.02860  1.8772 0.01878 0.099502 . 
#Residuals 66   1.00541 0.01523         0.66019            
#Total     69   1.52292                 1.00000  



###################
# Homogeneity of multivariate dispersion:

shallowDist <- vegdist(wisconsin(sqrt(shallowCPUE_spread3[,2:54])),"bray") # consider Hellinger transformation to deal with superabundant vs super rare species (see Legendre & Legendre)
#shallowDist = vegdist(wisconsin(sqrt(shallowCPUE_spread3[,2:54])),"jaccard") 


shallowTest_areas <- betadisper(shallowDist,shallowCPUE_spread3$Area) # this gives within-area variability (ie looks at temporal effects ... how different is each year from the mean across all years)
boxplot(shallowTest_areas) # indicates temporal turnover (beta diversity in time)
anova(shallowTest_areas) # p = 0.94   Shows that area of each hull in nmds is not significantly different ... ie no difference in temporal turnover between areas
#            Df   Sum Sq    Mean Sq F value Pr(>F)
# Groups      9 0.005072 0.00056356  0.3949 0.9357
# Residuals 130 0.185509 0.00142699 
# TukeyHSD(shallowTest_areas) # irrelevant


shallowTest_years <- betadisper(shallowDist,shallowCPUE_spread3$year) # within-year variability ... how different is each area from the mean across all areas?
boxplot(shallowTest_years)
anova(shallowTest_years) # p = 1 ... how does this add up with our signifant betwen-region result???
#            Df  Sum Sq   Mean Sq F value Pr(>F)
# Groups     13 0.00232 0.0001784  0.0378      1
# Residuals 126 0.59462 0.0047192
# TukeyHSD(shallowTest_years) # irrelevant



######################################################

# Deep areas:
adonis(wisconsin(sqrt(deepCPUE_spread3[,2:54])) ~ Area*year, data = deepCPUE_spread3, permutations = 200, method = "bray")
# Terms added sequentially (first to last)
# 
#          Df SumsOfSqs MeanSqs F.Model      R2   Pr(>F)   
#Area       1   0.31119 0.31119 16.2707 0.15524 0.004975 **
#year       1   0.41073 0.41073 21.4751 0.20490 0.004975 **
#Area:year  1   0.02035 0.02035  1.0642 0.01015 0.353234   
#Residuals 66   1.26230 0.01913         0.62971            
#Total     69   2.00457                 1.00000  

adonis(wisconsin(sqrt(deepCPUE_spread3[,2:54])) ~ Area, data = deepCPUE_spread3, permutations = 200, method = "bray")
#          Df SumsOfSqs  MeanSqs F.Model      R2   Pr(>F)   
#Area       1   0.31119 0.311189  12.496 0.15524 0.004975 **
#Residuals 68   1.69338 0.024903         0.84476            
#Total     69   2.00457                  1.00000 



###################
# Homogeneity of multivariate dispersion:

deepDist <- vegdist(wisconsin(sqrt(deepCPUE_spread3[,2:54])),"bray") # consider Hellinger transformation
#deepDist <- vegdist(wisconsin(sqrt(deepCPUE_spread3[,2:54])),"jaccard") 

deepTest <- betadisper(deepDist,deepCPUE_spread3$Area)
boxplot(deepTest)
anova(deepTest) # p = 0.89   
#           Df  Sum Sq    Mean Sq F value Pr(>F)
# Groups     4 0.00188 0.00046998   0.277 0.8919
# Residuals 65 0.11031 0.00169700  
# TukeyHSD(deepTest) # irrelevant


deepTest_years <- betadisper(deepDist,deepCPUE_spread3$year)
boxplot(deepTest_years)
anova(deepTest_years) # p = 0.95 ... how does this add up with our signifant betwen-region result???
#           Df   Sum Sq    Mean Sq F value Pr(>F)
# Groups    13 0.004998 0.00038448  0.4309 0.9515
# Residuals 56 0.049965 0.00089223 
# TukeyHSD(deepTest_years) # irrelevant

```



######################################################
######################################################
######################################################

#2. PCA
Methods: Used vegan package in R, function rda(). Raw CPUEs vary by many orders of magnitude (10^-7 to 10^2 or 10^-6 to 10^1 in shallow areas, 10^-5 to 10^2 in deep areas), therefore, to prevent the most abundant species from driving the result and allow all species to contribute equally to the result, cpues were standardized prior to analysis (Legendre & Legendre 1998) - means that PCs were extracted from correlation matrices, not dispersion matrices.

```{r,  include=FALSE, echo=FALSE, results='hide'}


# orgnzize the data for PCA:
# Shallow areas:
shallowCPUE_spread1 <- shallowCPUEArea2 %>%
  select(common.name, Mean.totalDensity, area, year) %>%
  filter(area != "Total") %>%
  spread(area, Mean.totalDensity) # make each area a column

shallowByYear_list <- split(shallowCPUE_spread1, f = shallowCPUE_spread1$year) # create a list of dataframes (one for each year)
shallowByYear_list1 <- lapply(shallowByYear_list, function(z){ row.names(z) <- z$common.name; z}) # create row names from common names
shallowByYear_list2 <- lapply(shallowByYear_list1, function(v) v[!(names(v) %in% c("common.name", "year"))]) # drop common name & year



# Deep areas:
deepCPUE_spread1 <- deepCPUEArea2 %>%
  select(common.name, Mean.totalDensity, area, year) %>%
  filter(area != "Total") %>%
  spread(area, Mean.totalDensity) # make each area a column

deepByYear_list <- split(deepCPUE_spread1, f = deepCPUE_spread1$year) # create a list of dataframes (one for each year)
deepByYear_list1 <- lapply(deepByYear_list, function(z){ row.names(z) <- z$common.name; z}) # create row names from common name
deepByYear_list2 <- lapply(deepByYear_list1, function(v) v[!(names(v) %in% c("common.name", "year"))]) # drop common name & year



#######################

# examine the data
cpueRangeShallow <- list(); mvnormShallow <- list()
for (i in seq_along(shallowByYear_list2)) {
  cpueRangeShallow[[i]] <- range(shallowByYear_list2[[i]]) # extract max & min cpue for each shallow area
  #mvnormShallow[[i]] <- mshapiro.test(as.matrix(shallowByYear_list2[[i]])) # assess multivariate normality
}
cpueRangeShallow; mvnormShallow
# p-values of multivariate shapiro-wilks test are all <<<<<<< 0.05, ie for each area, raw cpues are not multivariate-normal ... further suggestion that cpues should be standarized before analysis



# PCA for shallow areas
pcaShallow <- list()
for (i in seq_along(shallowByYear_list2)) {
  pcaShallow[[i]] <- rda(shallowByYear_list2[[i]], scale = TRUE)
}


# calculate proportion of total variance explained by PC1 & PC2:
pcVarShallow <- list()
for (i in seq_along(pcaShallow)) {
  pcVarShallow[[i]] <- pcaShallow[[i]]$CA[[1]] / pcaShallow[[i]]$CA$tot.chi
}

```

##Shallow areas
Area 10 is Total

% of total variance explained by PC1 & PC2:
```{r, echo=FALSE}
pcVarShallow1 <- data.frame(
  yr = c("1984", "1987", "1990", "1993", "1996", "1999", "2001", "2003", "2005", "2007", "2009", "2011", "2013", "2015"),
  pc1 <- do.call("rbind", lapply(pcVarShallow, "[[", 1)), # extract PC1
  pc2 <- do.call("rbind", lapply(pcVarShallow, "[[", 2))  # extract PC2
)
colnames <- c("Year", "PC1", "PC2"); names(pcVarShallow1 ) <- colnames
pcVarShallow1

```


Plots for shallow areas.
Note that I've removed arrows for species scores and did some wonky scaling of species scores to make them legible in the plots (divided each species score by species StDev - although their relative positions are preserved in the plots, this is technically incorrect because I've already standardized data prior to analysis). This will need to be changed if used in a publication - see .Rmd file for correct code.

```{r, echo=FALSE}
# plots:
yr = c("1984", "1987", "1990", "1993", "1996", "1999", "2001", "2003", "2005", "2007", "2009", "2011", "2013", "2015")
par(mfrow = c(1,1), mar=c(2,2,2,2))
pcaShallow1 <- list()
for (i in seq_along(pcaShallow)) {
  pcaShallow1[[i]] <- plot(pcaShallow[[i]], scaling = -1, main = paste("Shallow", yr[i])) # to add arrows for species scores, change this to biplot(...)
  # use of 'scaling = -1' is technically incorrect; for publication, change this to scaling = 3 or remove scaling = x 
}

```


***

```{r,  include=FALSE, echo=FALSE, results='hide'}
# Deep areas

# examine the data
cpueRangeDeep <- list(); mvnormDeep<- list()
for (i in seq_along(deepByYear_list2)) {
  cpueRangeDeep[[i]] <- range(deepByYear_list2[[i]]) # extract max & min cpue for each shallow area
  #mvnormDeep[[i]] <- mshapiro.test(as.matrix(deepByYear_list2[[i]])) # assess multivariate normality
}
cpueRangeDeep; mvnormDeep
# p-values of multivariate shapiro-wilks test are all <<<<<<< 0.05, ie for each area, raw cpues are not multivariate-normal ... further suggestion that cpues should be standarized before analysis


# PCA for deep areas
pcaDeep <- list()
for (i in seq_along(deepByYear_list2)) {
  pcaDeep[[i]] <- rda(deepByYear_list2[[i]], scale = TRUE)
}


# calculate proportion of total variance explained by PC1 & PC2:
pcVarDeep <- list()
for (i in seq_along(pcaDeep)) {
  pcVarDeep[[i]] <- pcaDeep[[i]]$CA[[1]] / pcaDeep[[i]]$CA$tot.chi
}

```
***
##Deep areas
Area 6 is Total

% of total variance explained by PC1 & PC2:
```{r, echo=FALSE}
pcVarDeep1 <- data.frame(
  yr = c("1984", "1987", "1990", "1993", "1996", "1999", "2001", "2003", "2005", "2007", "2009", "2011", "2013", "2015"),
  pc1.1 <- do.call("rbind", lapply(pcVarDeep, "[[", 1)), # extract PC1
  pc2.1 <- do.call("rbind", lapply(pcVarDeep, "[[", 2))  # extract PC2
)
colnames <- c("Year", "PC1", "PC2"); names(pcVarDeep1) <- colnames
pcVarDeep1

```

Plots for deep areas.
See note above about scaling of species scores.

```{r, echo=FALSE}

# Ordination plots:
par(mfrow = c(1,1), mar=c(2,2,2,2))
pcaDeep1 <- list()
for (i in seq_along(pcaDeep)) {
  pcaDeep1[[i]] <- plot(pcaDeep[[i]], scaling = -1, main = paste("Year", yr[i]))  # to add arrows for species scores, change this to biplot(...)
  # use of 'scaling = -1' is technically incorrect; for publication, change this to scaling = 3 or remove scaling = x
}

```


