---
title: "Groundfish multivariate analyses - Compare Years by Area"
author: "Colette Ward"
date: "May 5, 2016"
output: pdf_document
---

```{r, include=FALSE, echo=FALSE, results='hide'}
# Load packages
library(httr)
library(vegan)
library(plyr)
library(dplyr)
library(tidyr)

```

```{r,  include=FALSE, echo=FALSE, results='hide'}
# load & prep look-up table of common names
URL_common <- "https://drive.google.com/uc?export=download&id=0B1XbkXxdfD7ubzIzdURzanhfZnc" # this is "trawl_species_control_file.csv"
common_Get <- GET(URL_common)
common_1 <- content(common_Get, as='text')
common <- read.csv(file=textConnection(common_1),stringsAsFactors=FALSE,head=TRUE)

common1 <- common %>%
  select(database.name, common.name) %>%
  rename(Species = database.name)
for (i in 1:nrow(common1)) { # add common names for Sebastes 1 & 2
  if(common1$Species[i] == "Dusky.and.Dark.Rockfish") {common1$common.name[i] <- "Sebastes 1"}
  if(common1$Species[i] == "Rougheye.and.Blackspotted.Rockfish") {common1$common.name[i] <- "Sebastes 2"}
}
```


```{r,  include=FALSE, echo=FALSE, results='hide'}
# load & prepare Mean annual CPUE data for Shallow Areas:
URL_shallowCPUEArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uYzBOUFRtZklmX0U" # new data for shallow areas
shallowCPUEArea_Get <- GET(URL_shallowCPUEArea)
shallowCPUEArea_1 <- content(shallowCPUEArea_Get, as='text')
shallowCPUEArea <- read.csv(file=textConnection(shallowCPUEArea_1),stringsAsFactors=FALSE,head=TRUE)
#View(shallowCPUEArea)

shallowCPUEArea2 <- left_join(shallowCPUEArea, common1, by = "Species") # merge common names onto SPCPUEArea

shallowCPUE_spread <- shallowCPUEArea2 %>%
  select(common.name, Mean.totalDensity, area, year) %>%
  spread(common.name, Mean.totalDensity) %>% # make each species a column
  mutate(area = revalue(area, c("Total" = "10")), # recode Total for looping later
         area = as.numeric(area)) # convert to numeric class

shallowByArea_list <- split(shallowCPUE_spread, f = shallowCPUE_spread$area) # create a list of dataframes (one for each area; NB area 10 is Total)

shallowByArea_list1 <- lapply(shallowByArea_list, function(y){ row.names(y) <- y$year; y}) # create row names from year column
shallowByArea_list2 <- lapply(shallowByArea_list1, function(x) x[!(names(x) %in% c("area", "year"))]) # drop area & year

```


```{r,  include=FALSE, echo=FALSE, results='hide'}
# load & prepare mean annual CPUE for Deep areas:
URL_deepCPUEArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uVF9VWnNPX3Z3S3c"
deepCPUEArea_Get <- GET(URL_deepCPUEArea)
deepCPUEArea_1 <- content(deepCPUEArea_Get, as='text')
deepCPUEArea <- read.csv(file=textConnection(deepCPUEArea_1),stringsAsFactors=FALSE,head=TRUE)
#View(deepCPUEArea)

deepCPUEArea2 <- left_join(deepCPUEArea, common1, by = "Species") # merge common names onto SPCPUEArea

deepCPUE_spread <- deepCPUEArea2 %>%
  select(common.name, Mean.totalDensity, area, year) %>%
  spread(common.name, Mean.totalDensity) %>% # make each species a column
  mutate(area = revalue(area, c("Total" = "10")), # recode Total for looping later
         area = as.numeric(area)) # convert to numeric class

deepByArea_list <- split(deepCPUE_spread, f = deepCPUE_spread$area) # create a list of dataframes (one for each area; NB area 10 is Total)

deepByArea_list1 <- lapply(deepByArea_list, function(y){ row.names(y) <- y$year; y}) # create row names from year column
deepByArea_list2 <- lapply(deepByArea_list1, function(x) x[!(names(x) %in% c("area", "year"))]) # drop area & year

```

######################################################
######################################################
######################################################

1. nMDS

Transformation is Wisconsin-style double transformation (normalizes taxa to % abundance, then normalizes abundances to the maximum for each species) of sqrt-transformed data.

Sorry the text is so small in the ordination plots, need small text for legibility because there are so many species. 
Enlarge the output pdf to see species names.


```{r,  include=FALSE, echo=FALSE, results='hide'}

# nMDS for Shallow Areas:
ordShallow <- list()
for (i in seq_along(shallowByArea_list2)) {
  ordShallow[[i]] <- metaMDS(shallowByArea_list2[[i]], distance='bray', trymax=1000)
}


# examine Shepard's stressplots:
# goodness-of-fit test: plots ordination distances against community dissimilarities
par (mfrow = c(3,4), pty="m") 
stressplotsShallow <- list()
for (i in seq_along(ordShallow)) {
  stressplotsShallow[[i]] <- stressplot(ordShallow[[i]], main = paste("Area",i))
}


# examine goodness of fit plots:
# size represents goodness of fit (bigger = worse fit)
par (mfrow = c(3,4), pty="m") 
goodnessShallow <- list()
for (i in seq_along(ordShallow)) {
  goodnessShallow[[i]] <- plot (ordShallow[[i]], display = 'sites', type = 't', main = paste("Area",i))
  points (ordShallow[[i]], display = 'sites', cex = goodness (ordShallow[[i]])*200)
}

```

#Shallow areas
Area 10 is Total

```{r, echo=FALSE}
# Ordination plots:
par(mfrow = c(2,2), mar=c(2,2,2,2))
ordPlotsShallow1 <- list()
for (i in seq_along(ordShallow)) {
  ordPlotsShallow1[[i]] <- plot(ordShallow[[i]], type = "n", main = paste("Shallow Area",i))
  text(ordShallow[[i]], display = "spec", cex = 0.25, col="red")
  text(ordShallow[[i]], display = "sites", cex=0.7, col="blue")
}
```

***


```{r,  include=FALSE, echo=FALSE, results='hide'}

# nMDS for Deep Areas:
ordDeep <- list()
for (i in seq_along(deepByArea_list2)) {
  ordDeep[[i]] <- metaMDS(deepByArea_list2[[i]], distance='bray', trymax=1000)
}


# examine Shepard's stressplots:
# goodness-of-fit test: plots ordination distances against community dissimilarities
par (mfrow = c(2,3), pty="m") 
stressplotsDeep <- list()
for (i in seq_along(ordDeep)) {
  stressplotsDeep[[i]] <- stressplot(ordDeep[[i]], main = paste("Area",i))
}


# examine goodness of fit plots:
# size represents goodness of fit (bigger = worse fit)
par (mfrow = c(2,3), pty="m") 
goodnessDeep <- list()
for (i in seq_along(ordDeep)) {
  goodnessDeep[[i]] <- plot (ordDeep[[i]], display = 'sites', type = 't', main = paste("Area",i))
  points (ordDeep[[i]], display = 'sites', cex = goodness (ordDeep[[i]])*200)
}
```

#Deep areas
Area 6 is Total

```{r, echo=FALSE}
# Ordination plots:
# NB Area 6 is total of our 5 areas
par(mfrow = c(2,2), mar=c(2,2,2,2))
ordPlotsDeep1 <- list()
for (i in seq_along(ordDeep)) {
  ordPlotsDeep1[[i]] <- plot(ordDeep[[i]], type = "n", main = paste("Deep Area",i))
  text(ordDeep[[i]], display = "spec", cex = 0.25, col="red")
  text(ordDeep[[i]], display = "sites", cex=0.7, col="blue")
}
```


######################################################
######################################################
######################################################

2. PCA

```{r,  include=FALSE, echo=FALSE, results='hide'}

# PCA for shallow areas
pcaShallow <- list()
for (i in seq_along(shallowByArea_list2)) {
  pcaShallow[[i]] <- rda(shallowByArea_list2[[i]], scale = TRUE)
}

```

#Shallow areas
Area 10 is Total

```{r, echo=FALSE}

# Ordination plots:
par(mfrow = c(2,2), mar=c(2,2,2,2))
pcaShallow1 <- list()
for (i in seq_along(pcaShallow)) {
  pcaShallow1[[i]] <- plot(pcaShallow[[i]], scaling = 3, main = paste("Shallow Area",i))
}
```

***

```{r,  include=FALSE, echo=FALSE, results='hide'}

# PCA for deep areas
pcaDeep <- list()
for (i in seq_along(deepByArea_list2)) {
  pcaDeep[[i]] <- rda(deepByArea_list2[[i]], scale = TRUE)
}
```

#Deep areas
Area 6 is Total

```{r, echo=FALSE}

# Ordination plots:
par(mfrow = c(2,2), mar=c(2,2,2,2))
pcaDeep1 <- list()
for (i in seq_along(pcaDeep)) {
  pcaDeep1[[i]] <- plot(pcaDeep[[i]], scaling = 3, main = paste("Deep Area",i))
}
```