---
title: "Groundfish multivariate analyses - Compare Years by Area"
author: "Colette Ward"
date: "May 5, 2016"
output: pdf_document
---

```{r, include=FALSE, echo=FALSE, results='hide'}
# Load packages
library(httr)
library(vegan)
library(mvnormtest)
library(plyr)
library(dplyr)
library(tidyr)

```

```{r,  include=FALSE, echo=FALSE, results='hide'}
# load & prep look-up table of common names
URL_common <- "https://drive.google.com/uc?export=download&id=0B1XbkXxdfD7ubzIzdURzanhfZnc" # this is "trawl_species_control_file.csv"
common_Get <- GET(URL_common)
common_1 <- content(common_Get, as='text')
common <- read.csv(file=textConnection(common_1),stringsAsFactors=FALSE,head=TRUE)

common1 <- common %>%
  select(database.name, common.name) %>%
  rename(Species = database.name)
for (i in 1:nrow(common1)) { # add common names for Sebastes 1 & 2
  if(common1$Species[i] == "Dusky.and.Dark.Rockfish") {common1$common.name[i] <- "Sebastes 1"}
  if(common1$Species[i] == "Rougheye.and.Blackspotted.Rockfish") {common1$common.name[i] <- "Sebastes 2"}
}
```


```{r,  include=FALSE, echo=FALSE, results='hide'}
# load & prepare Mean annual CPUE data for Shallow Areas:
URL_shallowCPUEArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uYzBOUFRtZklmX0U" # new data for shallow areas
shallowCPUEArea_Get <- GET(URL_shallowCPUEArea)
shallowCPUEArea_1 <- content(shallowCPUEArea_Get, as='text')
shallowCPUEArea <- read.csv(file=textConnection(shallowCPUEArea_1),stringsAsFactors=FALSE,head=TRUE)
#View(shallowCPUEArea)

shallowCPUEArea2 <- left_join(shallowCPUEArea, common1, by = "Species") # merge common names onto SPCPUEArea

shallowCPUE_spread <- shallowCPUEArea2 %>%
  select(common.name, Mean.totalDensity, area, year) %>%
  spread(common.name, Mean.totalDensity) %>% # make each species a column
  mutate(area = revalue(area, c("Total" = "10")), # recode Total for looping later
         area = as.numeric(area)) # convert to numeric class

shallowByArea_list <- split(shallowCPUE_spread, f = shallowCPUE_spread$area) # create a list of dataframes (one for each area; NB area 10 is Total)

shallowByArea_list1 <- lapply(shallowByArea_list, function(y){ row.names(y) <- y$year; y}) # create row names from year column
shallowByArea_list2 <- lapply(shallowByArea_list1, function(x) x[!(names(x) %in% c("area", "year"))]) # drop area & year

```


```{r,  include=FALSE, echo=FALSE, results='hide'}
# load & prepare mean annual CPUE for Deep areas:
URL_deepCPUEArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uVF9VWnNPX3Z3S3c"
deepCPUEArea_Get <- GET(URL_deepCPUEArea)
deepCPUEArea_1 <- content(deepCPUEArea_Get, as='text')
deepCPUEArea <- read.csv(file=textConnection(deepCPUEArea_1),stringsAsFactors=FALSE,head=TRUE)
#View(deepCPUEArea)

deepCPUEArea2 <- left_join(deepCPUEArea, common1, by = "Species") # merge common names onto SPCPUEArea

deepCPUE_spread <- deepCPUEArea2 %>%
  select(common.name, Mean.totalDensity, area, year) %>%
  spread(common.name, Mean.totalDensity) %>% # make each species a column
  mutate(area = revalue(area, c("Total" = "10")), # recode Total for looping later
         area = as.numeric(area)) # convert to numeric class

deepByArea_list <- split(deepCPUE_spread, f = deepCPUE_spread$area) # create a list of dataframes (one for each area; NB area 10 is Total)

deepByArea_list1 <- lapply(deepByArea_list, function(y){ row.names(y) <- y$year; y}) # create row names from year column
deepByArea_list2 <- lapply(deepByArea_list1, function(x) x[!(names(x) %in% c("area", "year"))]) # drop area & year

```

######################################################
######################################################
######################################################

#1. nMDS

Transformation is Wisconsin-style double transformation (normalizes taxa to % abundance, then normalizes abundances to the maximum for each species) of sqrt-transformed data.

**Sorry the text is so small. It's necessary for legibility because there are so many species.** 
**Enlarge the output pdf to see species names.**


```{r,  include=FALSE, echo=FALSE, results='hide'}

# nMDS for Shallow Areas:
ordShallow <- list()
for (i in seq_along(shallowByArea_list2)) {
  ordShallow[[i]] <- metaMDS(shallowByArea_list2[[i]], distance='bray', trymax=1000)
}


# examine Shepard's stressplots:
# goodness-of-fit test: plots ordination distances against community dissimilarities
par (mfrow = c(3,4), pty="m") 
stressplotsShallow <- list()
for (i in seq_along(ordShallow)) {
  stressplotsShallow[[i]] <- stressplot(ordShallow[[i]], main = paste("Area",i))
}


# examine goodness of fit plots:
# size represents goodness of fit (bigger = worse fit)
par (mfrow = c(3,4), pty="m") 
goodnessShallow <- list()
for (i in seq_along(ordShallow)) {
  goodnessShallow[[i]] <- plot (ordShallow[[i]], display = 'sites', type = 't', main = paste("Area",i))
  points (ordShallow[[i]], display = 'sites', cex = goodness (ordShallow[[i]])*200)
}

```

##Shallow areas
Area 10 is Total

```{r, echo=FALSE}
# Ordination plots:
par(mfrow = c(1,2), mar=c(2,2,2,2), pty = "s")
ordPlotsShallow1 <- list()
for (i in seq_along(ordShallow)) {
  ordPlotsShallow1[[i]] <- plot(ordShallow[[i]], type = "n", main = paste("Shallow Area",i))
  text(ordShallow[[i]], display = "spec", cex = 0.25, col="red")
  text(ordShallow[[i]], display = "sites", cex=0.7, col="blue")
}
```

***


```{r,  include=FALSE, echo=FALSE, results='hide'}

# nMDS for Deep Areas:
ordDeep <- list()
for (i in seq_along(deepByArea_list2)) {
  ordDeep[[i]] <- metaMDS(deepByArea_list2[[i]], distance='bray', trymax=1000)
}


# examine Shepard's stressplots:
# goodness-of-fit test: plots ordination distances against community dissimilarities
par (mfrow = c(2,3), pty="m") 
stressplotsDeep <- list()
for (i in seq_along(ordDeep)) {
  stressplotsDeep[[i]] <- stressplot(ordDeep[[i]], main = paste("Area",i))
}


# examine goodness of fit plots:
# size represents goodness of fit (bigger = worse fit)
par (mfrow = c(2,3), pty="m") 
goodnessDeep <- list()
for (i in seq_along(ordDeep)) {
  goodnessDeep[[i]] <- plot (ordDeep[[i]], display = 'sites', type = 't', main = paste("Area",i))
  points (ordDeep[[i]], display = 'sites', cex = goodness (ordDeep[[i]])*200)
}
```

##Deep areas
Area 6 is Total

```{r, echo=FALSE}
# Ordination plots:
# NB Area 6 is total of our 5 areas
par(mfrow = c(2,2), mar=c(2,2,2,2))
ordPlotsDeep1 <- list()
for (i in seq_along(ordDeep)) {
  ordPlotsDeep1[[i]] <- plot(ordDeep[[i]], type = "n", main = paste("Deep Area",i))
  text(ordDeep[[i]], display = "spec", cex = 0.25, col="red")
  text(ordDeep[[i]], display = "sites", cex=0.7, col="blue")
}
```
***

######################################################
######################################################
######################################################

#2. PCA
Methods: Used vegan package in R, function rda(). Raw CPUEs vary by many orders of magnitude (10^-7 to 10^2 in shallow areas, 10^-5 to 10^2 in deep areas), therefore, to prevent the most abundant species from driving the result and allow all species to contribute equally to the result, cpues were standardized prior to analysis (Legendre & Legendre 1998) - means that PCs were extracted from correlation matrices, not dispersion matrices.

```{r,  include=FALSE, echo=FALSE, results='hide'}

# examine the data
cpueRangeShallow <- list(); mvnormShallow <- list()
for (i in seq_along(shallowByArea_list2)) {
  cpueRangeShallow[[i]] <- range(shallowByArea_list2[[i]]) # extract max & min cpue for each shallow area
  mvnormShallow[[i]] <- mshapiro.test(as.matrix(shallowByArea_list2[[i]])) # assess multivariate normality
}
cpueRangeShallow; mvnormShallow
# p-values of multivariate shapiro-wilks test are all <<<<<<< 0.05, ie for each area, raw cpues are not multivariate-normal ... further suggestion that cpues should be standarized before analysis



# PCA for shallow areas
pcaShallow <- list()
for (i in seq_along(shallowByArea_list2)) {
  pcaShallow[[i]] <- rda(shallowByArea_list2[[i]], scale = TRUE)
}


# calculate proportion of total variance explained by PC1 & PC2:
pcVarShallow <- list()
for (i in seq_along(pcaShallow)) {
  pcVarShallow[[i]] <- pcaShallow[[i]]$CA[[1]] / pcaShallow[[i]]$CA$tot.chi
}

```

##Shallow areas
Area 10 is Total

% of total variance explained by PC1 & PC2 (Area 10 = Total):
```{r, echo=FALSE}
pcVarShallow1 <- data.frame(
  area <- c(1:10),
  pc1 <- do.call("rbind", lapply(pcVarShallow, "[[", 1)), # extract PC1
  pc2 <- do.call("rbind", lapply(pcVarShallow, "[[", 2))  # extract PC2
)
colnames <- c("Area", "PC1", "PC2"); names(pcVarShallow1 ) <- colnames
pcVarShallow1

```


Plots for shallow areas.
Note that I've removed arrows for species scores and did some wonky scaling of species scores to make them legible in the plots (divided each species score by species StDev - although their relative positions are preserved in the plots, this is technically incorrect because I've already standardized data prior to analysis). This will need to be changed if used in a publication - see .Rmd file for correct code.

```{r, echo=FALSE}
# Ordination plots:
par(mfrow = c(1,1), mar=c(2,2,2,2))
pcaShallow1 <- list()
for (i in seq_along(pcaShallow)) {
  pcaShallow1[[i]] <- plot(pcaShallow[[i]], scaling = -1, main = paste("Shallow Area",i)) # to add arrows for species scores, change this to biplot(...)
  # use of 'scaling = -1' is technically incorrect; for publication, change this to scaling = 3 or remove scaling = x 
}

```


***

```{r,  include=FALSE, echo=FALSE, results='hide'}
# Deep areas

# examine the data
cpueRangeDeep <- list(); mvnormDeep<- list()
for (i in seq_along(deepByArea_list2)) {
  cpueRangeDeep[[i]] <- range(deepByArea_list2[[i]]) # extract max & min cpue for each shallow area
  mvnormDeep[[i]] <- mshapiro.test(as.matrix(deepByArea_list2[[i]])) # assess multivariate normality
}
cpueRangeDeep; mvnormDeep
# p-values of multivariate shapiro-wilks test are all <<<<<<< 0.05, ie for each area, raw cpues are not multivariate-normal ... further suggestion that cpues should be standarized before analysis


# PCA for deep areas
pcaDeep <- list()
for (i in seq_along(deepByArea_list2)) {
  pcaDeep[[i]] <- rda(deepByArea_list2[[i]], scale = TRUE)
}


# calculate proportion of total variance explained by PC1 & PC2:
pcVarDeep <- list()
for (i in seq_along(pcaDeep)) {
  pcVarDeep[[i]] <- pcaDeep[[i]]$CA[[1]] / pcaDeep[[i]]$CA$tot.chi
}

```
***
##Deep areas
Area 6 is Total

% of total variance explained by PC1 & PC2 (Area 6 = Total):
```{r, echo=FALSE}
pcVarDeep1 <- data.frame(
  area <- c(1:6),
  pc1.1 <- do.call("rbind", lapply(pcVarDeep, "[[", 1)), # extract PC1
  pc2.1 <- do.call("rbind", lapply(pcVarDeep, "[[", 2))  # extract PC2
)
colnames <- c("Area", "PC1", "PC2"); names(pcVarDeep1) <- colnames
pcVarDeep1

```

Plots for deep areas.
See note above about scaling of species scores.

```{r, echo=FALSE}

# Ordination plots:
par(mfrow = c(1,1), mar=c(2,2,2,2))
pcaDeep1 <- list()
for (i in seq_along(pcaDeep)) {
  pcaDeep1[[i]] <- plot(pcaDeep[[i]], scaling = -1, main = paste("Deep Area",i))  # to add arrows for species scores, change this to biplot(...)
  # use of 'scaling = -1' is technically incorrect; for publication, change this to scaling = 3 or remove scaling = x
}

#plot(pcaDeep[[1]], scaling = -1)
#biplot(pcaDeep[[1]], scaling = -1)
# scaling = -1: The negative values mean that each species score is divided by its standard deviation so that abundant and scarce species will be approximately as far away from the origin.
```