---
title: "portfolio_plots.rmd"
author: "Ole Shelton"
date: "March 8, 2016"
output: pdf_document
---

This is a file for trying out alternative plots of portfolio metrics for groundfish.

Here are a bunch of plots for looking at trend and variability in different groups for the period 1990 to present

```{r load, message=FALSE, echo=FALSE, warning=FALSE}
library(extrafont)
#font_import() #only do this one time - it takes a while
loadfonts(quiet = T)
#windowsFonts(Times=windowsFont("TT Times New Roman"))

library(ggplot2)
library(maps)
library(mapdata)
library(mapproj)
library(dplyr)
library(tidyr)

# Go get the needed data:

meanCPUE <- read.csv("All_sp_index_meanCPUEByArea.csv") # load data
meanCPUE <- subset(meanCPUE,area!='Total') #remove total field
meanCPUE <- droplevels(meanCPUE) 
meanCPUE$area <- as.numeric(levels(meanCPUE$area))[meanCPUE$area]#use numeric/factor conventions
meanCPUE$Area <- factor(meanCPUE$area)
meanCPUE$vari <- meanCPUE$SD.totalDensity^2 #calculate variance
#meanCPUE$Species <- gsub("[.]","",meanCPUE$Species) #quit changing how names are spelled damnit!

discrete_areas <- read.csv("./goaTrawl/Output Data/goa_discrete_areas_for_comparison(50_to_150m).csv")
#head(discrete_areas)

trawl_species <- read.csv("./goaTrawl/Output Data/trawl_species_control_file.csv")
s <- trawl_species$diet
s <- as.data.frame(s) %>% separate(s, into = paste("diet", 1:2, sep = ""))
trawl_species <- cbind(trawl_species,s)

### Merge in the guild identifiers, fish habit, and other functional categories into the meanCPUE data.frame
temp <- trawl_species[,c("database.name","fish.invert","pelagic.benthic","total.biomass.fish","guild","diet1","diet2")]
colnames(temp)[1] <- "Species"

meanCPUE <- merge(meanCPUE,temp)
```

```{r, echo=FALSE}
weight.var <- function(w){
  return(sum(w)^(-1))
}
space_plot <- function(dat,NAME){
    par(mfrow=c(3,2),mar=c(3,4,1,0.5))
    AT <- dat$Area  
    plot(1:5,1:5,type="n",axes=F,xlab="",ylab="")
    text(1.5,4,NAME,cex=3,pos=4)
    y.lim <- c(min(dat$Trend - dat$SE.trend),max(dat$Trend + dat$SE.trend))
    plot(Trend~Area,data=dat,ylim=y.lim,axes=F,pch=21,bg=1)
    abline(h=0,lty=2)
    arrows(x0=dat$Area,x1=dat$Area,
          y0=dat$Trend + dat$SE,y1= dat$Trend - dat$SE,
          length=0,lwd=1.5)
      axis(1,at=AT); box(bty="o")
      axis(2,las=2)
      title(xlab="Area",line=2)
    plot(Grand.w.mean~Area,data=dat,type="b",axes=F,xlab="",pch=21,bg=1)
      axis(1,at=AT); box(bty="o")
      axis(2,las=2)
      title(xlab="Area",line=2)
    plot(SD~Area,data=dat,type="b",axes=F,ylim= c(0,max(dat$SD)),xlab="",pch=21,bg=1)
      axis(1,at=AT); box(bty="o")
      axis(2,las=2)
      title(xlab="Area",line=2)
    plot(CV~Area,data=dat,type="b",axes=F,xlab="",pch=21,bg=1)
      axis(1,at=AT); box(bty="o")
      axis(2,las=2)
      title(xlab="Area",line=2)
    plot(Var~Grand.w.mean,data=dat,axes=F,xlab="",pch=21,bg=1)
      axis(1); box(bty="o")
      axis(2,las=2)
      title(xlab="W.mean",line=2)
}
```


```{r map, cache=FALSE, echo=FALSE, warning=FALSE}
ak<-map_data('worldHires','USA:Alaska')
akmap<-ggplot()+geom_polygon(data=ak,aes(long,lat,group=group),fill=8,color="black")+
    theme(panel.background=element_rect(fill='aliceblue'))+
    xlab(expression(paste(Longitude^o,~'W')))+
    ylab(expression(paste(Latitude^o,~'N')))+
    coord_map(xlim = c(-160, -143),ylim = c(54, 61))
```

Map of area locations

```{r, echo=FALSE, cache=FALSE}
discrete_areas$area <- discrete_areas$Area
discrete_areas$Area <- factor(discrete_areas$area)
akmap+geom_point(data=discrete_areas,aes(LONGITUDE,LATITUDE,color=Area))
```


Plot time series of a few community metrics

```{r, echo=FALSE}

time_plot <- function(dat,NAME){
  AREA <- as.numeric(as.character(unique(dat$Area)))
  YEAR <- sort(unique(dat$Year))
  y.lim <- c(min(dat$Mean.totalDensity),max(dat$Mean.totalDensity))
  for(i in 1:length(AREA)){
    if(AREA[i] !=1){par(new=T)}
    if(AREA[i] >= 3 & AREA[i] <= 5 ){
      plot(Mean.totalDensity~Year,data=dat[dat$Area ==AREA[i],],
            type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
    }
    if(AREA[i] < 3 | AREA[i] > 5){
      plot(Mean.totalDensity~Year,data=dat[dat$Area ==AREA[i],],
            type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
    }
  }
    axis(1,at=YEAR,las=2)
    axis(2,las=2)
    box(bty="o",lwd=2)
    title(ylab=expression("Biomass (kg ha"^"-1"*")"),line=2.5)
    mtext(NAME,adj=0)
    abline(v=1989,lty=2,lwd=2)
}

```


Total Biomass:
(RED denotes areas 3-5)

```{r, echo=FALSE}
# TOTAL BIOMASS
  dat   <- aggregate(meanCPUE[meanCPUE$fish.invert=="fish",c("Mean.totalDensity","vari")],
            by=list(Area=meanCPUE$Area[meanCPUE$fish.invert=="fish"],Year=meanCPUE$year[meanCPUE$fish.invert=="fish"]),sum)
  dat$inv.var <- dat$vari^(-1)
  dat$x.bar <- dat$Mean.totalDensity * dat$inv.var

  par(mfrow=c(1,1),mar=c(3,5,2,2))
  time_plot(dat,NAME="Total Biomass")
```


```{r, echo=FALSE}
# GUILDS
  par(mfrow=c(2,2),mar=c(3,4,2,0.5))

  GUILD <- unique(meanCPUE$guild)
  for(i in 1:length(GUILD)){
    temp <- meanCPUE[meanCPUE$guild == GUILD[i],]
    dat   <- aggregate(temp[,c("Mean.totalDensity","vari")],by=list(Area=temp$Area,Year=temp$year),sum)
    dat$inv.var <- dat$vari^(-1)
    dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
    time_plot(dat,NAME=paste("Guild",GUILD[i]))
  }

```


```{r, echo=FALSE}
# Lifestyle
  par(mfrow=c(2,2),mar=c(3,4,2,0.5))

  LIFESTYLE <- unique(meanCPUE$pelagic.benthic)
  for(i in 1:length(LIFESTYLE)){
    temp <- meanCPUE[meanCPUE$pelagic.benthic == LIFESTYLE[i],]
    dat   <- aggregate(temp[,c("Mean.totalDensity","vari")],by=list(Area=temp$Area,Year=temp$year),sum)
    dat$inv.var <- dat$vari^(-1)
    dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
    time_plot(dat,NAME=paste(LIFESTYLE[i]))
  }
```

Total Biomass (fish only)

```{r, echo=FALSE}

# TOTAL BIOMASS
  dat   <- aggregate(meanCPUE[meanCPUE$fish.invert=="fish",c("Mean.totalDensity","vari")],
            by=list(Area=meanCPUE$Area[meanCPUE$fish.invert=="fish"],Year=meanCPUE$year[meanCPUE$fish.invert=="fish"]),sum)
  dat$inv.var <- dat$vari^(-1)
  dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
    ### Calculate least squares trend in each area and portfolio metrics.
 AREA <- unique(dat$Area)
  vals <- NULL
  for(j in 1:length(AREA)){
    A <- lm(Mean.totalDensity~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,],weights = inv.var )
    B <- lm(Mean.totalDensity~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,],weights = inv.var )
    vals <- rbind(vals,c(AREA[j],
                            coef(summary(A))["Year",c("Estimate","Std. Error")],
                            coef(summary(B))[1,c("Estimate","Std. Error")],
                            var(A$residuals),
                            sd(A$residuals)))
                         
  }
  vals <- data.frame(vals)
  colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
  vals$CV  <- vals$SD / vals$Grand.w.mean 
  
  dat.2 <- vals
  
  space_plot(dat=dat.2,NAME=paste("Total Biomass"))

```  
  
 **By Feeding Guild**
 
  * A = Apex predators (Piscivores)
  * B = Benthic Foragers
  * P = Pelagic Foragers
  * E = Invertebrates (benthic foragers)

```{r, echo=FALSE}  
# Aggregate into groups by guild and make plots
GUILD <- unique(meanCPUE$guild)
for(i in 1:length(GUILD)){
  temp <- meanCPUE[meanCPUE$guild == GUILD[i],]
  dat   <- aggregate(temp[,c("Mean.totalDensity","vari")],by=list(Area=temp$Area,Year=temp$year),sum)
  dat$inv.var <- dat$vari^(-1)
  dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
    ### Calculate least squares trend in each area and portfolio metrics.
 AREA <- unique(dat$Area)
  vals <- NULL
  for(j in 1:length(AREA)){
    A <- lm(Mean.totalDensity~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,],weights = inv.var )
    B <- lm(Mean.totalDensity~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,],weights = inv.var )
    vals <- rbind(vals,c(AREA[j],
                            coef(summary(A))["Year",c("Estimate","Std. Error")],
                            coef(summary(B))[1,c("Estimate","Std. Error")],
                            var(A$residuals),
                            sd(A$residuals)))
                         
  }
  vals <- data.frame(vals)
  colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
  vals$CV  <- vals$SD / vals$Grand.w.mean 
  
  dat.2 <- vals
  
  space_plot(dat=dat.2,NAME=paste("Guild ",GUILD[i]))
}
```


 **By Benthic vs. Pelagic**
 
```{r, echo=FALSE}  
# Aggregate into groups by guild and make plots
LIFESTYLE <- unique(meanCPUE$pelagic.benthic)
for(i in 1:length(LIFESTYLE)){
  temp <- meanCPUE[meanCPUE$pelagic.benthic == LIFESTYLE[i],]
  dat   <- aggregate(temp[,c("Mean.totalDensity","vari")],by=list(Area=temp$Area,Year=temp$year),sum)
  dat$inv.var <- dat$vari^(-1)
  dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
     ### Calculate least squares trend in each area and portfolio metrics.
 AREA <- unique(dat$Area)
  vals <- NULL
  for(j in 1:length(AREA)){
    A <- lm(Mean.totalDensity~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,],weights = inv.var )
    B <- lm(Mean.totalDensity~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,],weights = inv.var )
    vals <- rbind(vals,c(AREA[j],
                            coef(summary(A))["Year",c("Estimate","Std. Error")],
                            coef(summary(B))[1,c("Estimate","Std. Error")],
                            var(A$residuals),
                            sd(A$residuals)))
                         
  }
  vals <- data.frame(vals)
  colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
  vals$CV  <- vals$SD / vals$Grand.w.mean 
  
  dat.2 <- vals
  
  space_plot(dat=dat.2,NAME=paste(LIFESTYLE[i]))
}
```

 **By DIET**
 
* I = invertebrate prey
* F = fish prey
* G = generalist (Generalist if consumes both Invert & Fish prey (threshold = 20% diet by weight))

* B = demersal prey
* C = pelagic prey
  
```{r, echo=FALSE}  
# Aggregate into groups by guild and make plots
DIET <- unique(meanCPUE$diet1)
for(i in 1:length(DIET)){
  temp <- meanCPUE[meanCPUE$diet1 == DIET[i],]
  dat   <- aggregate(temp[,c("Mean.totalDensity","vari")],by=list(Area=temp$Area,Year=temp$year),sum)
  dat$inv.var <- dat$vari^(-1)
  dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
     ### Calculate least squares trend in each area and portfolio metrics.
 AREA <- unique(dat$Area)
  vals <- NULL
  for(j in 1:length(AREA)){
    A <- lm(Mean.totalDensity~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,],weights = inv.var )
    B <- lm(Mean.totalDensity~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,],weights = inv.var )
    vals <- rbind(vals,c(AREA[j],
                            coef(summary(A))["Year",c("Estimate","Std. Error")],
                            coef(summary(B))[1,c("Estimate","Std. Error")],
                            var(A$residuals),
                            sd(A$residuals)))
                         
  }
  vals <- data.frame(vals)
  colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
  vals$CV  <- vals$SD / vals$Grand.w.mean 
  
  dat.2 <- vals
  
  space_plot(dat=dat.2,NAME=paste("Diet type" ,DIET[i]))
}
```


```{r, echo=FALSE}  
# Aggregate into groups by guild and make plots
DIET <- unique(meanCPUE$diet2)
  DIET <- DIET[is.na(DIET)==F]
for(i in 1:length(DIET)){
  temp <- meanCPUE[meanCPUE$diet2 == DIET[i],]
  dat   <- aggregate(temp[,c("Mean.totalDensity","vari")],by=list(Area=temp$Area,Year=temp$year),sum)
  dat$inv.var <- dat$vari^(-1)
  dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
     ### Calculate least squares trend in each area and portfolio metrics.
 AREA <- unique(dat$Area)
  vals <- NULL
  for(j in 1:length(AREA)){
    A <- lm(Mean.totalDensity~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,],weights = inv.var )
    B <- lm(Mean.totalDensity~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,],weights = inv.var )
    vals <- rbind(vals,c(AREA[j],
                            coef(summary(A))["Year",c("Estimate","Std. Error")],
                            coef(summary(B))[1,c("Estimate","Std. Error")],
                            var(A$residuals),
                            sd(A$residuals)))
                         
  }
  vals <- data.frame(vals)
  colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
  vals$CV  <- vals$SD / vals$Grand.w.mean 
  
  dat.2 <- vals
  
  space_plot(dat=dat.2,NAME=paste("Diet type" ,DIET[i]))
}
```

**Diversity Metics**
```{r, echo=FALSE}
#Calcuate metrics

libary(vegan)
libary(dplyr)

meanCPUE$Area <- meanCPUE$area
meanCPUE$Year<-meanCPUE$year

DivMetrics <- meanCPUE %>%
  group_by(Area, Year) %>% 
  summarise(SW_Div = diversity(Mean.totalDensity, index="shannon"), Eff_Num_Sp = exp(SW_Div),
            simp_Div = diversity(Mean.totalDensity, index="simpson"), 
            invsimp = diversity(Mean.totalDensity,index="inv")) %>%
  ungroup() %>%
  print(width = Inf)

dat<-DivMetrics
AREA <- as.numeric(as.character(unique(dat$Area)))
YEAR <- sort(unique(dat$Year))
```
Species Richness
```{r, echo=FALSE}
Plot time series by area
y.lim <- c(min(dat$Eff_Num_Sp),max(dat$Eff_Num_Sp))
for(i in 1:length(AREA)){
  if(AREA[i] !=1){par(new=T)}
  if(AREA[i] >= 3 & AREA[i] <= 5 ){
    plot(Eff_Num_Sp~Year,data=dat[dat$Area==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
  }
  if(AREA[i] < 3 | AREA[i] > 5){
    plot(Eff_Num_Sp~Year,data=dat[dat$Area ==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
  }
}
axis(1,at=YEAR,las=2)
axis(2,las=2)
box(bty="o",lwd=2)
title(ylab="Species Richness",line=2.5)
#mtext(NAME,adj=0)
abline(v=1989,lty=2,lwd=2)

#Calculate least squares trend in each area and portfolio metrics
vals <- NULL
for(j in 1:length(AREA)){
  A <- lm(Eff_Num_Sp~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  B <- lm(Eff_Num_Sp~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  vals <- rbind(vals,c(AREA[j],
                       coef(summary(A))["Year",c("Estimate","Std. Error")],
                       coef(summary(B))[1,c("Estimate","Std. Error")],
                       var(A$residuals),
                       sd(A$residuals)))
  
}
vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(dat=dat.2,NAME='Species Richness')
```

Shannon Wiener Index
```{r, echo=FALSE}
#Plot time series by area
y.lim <- c(min(dat$SW_Div),max(dat$SW_Div))
for(i in 1:length(AREA)){
  if(AREA[i] !=1){par(new=T)}
  if(AREA[i] >= 3 & AREA[i] <= 5 ){
    plot(SW_Div~Year,data=dat[dat$Area==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
  }
  if(AREA[i] < 3 | AREA[i] > 5){
    plot(SW_Div~Year,data=dat[dat$Area ==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
  }
}
axis(1,at=YEAR,las=2)
axis(2,las=2)
box(bty="o",lwd=2)
title(ylab="Shannon Wiener Index",line=2.5)
#mtext(NAME,adj=0)
abline(v=1989,lty=2,lwd=2)

#Calculate least squares trend in each area and portfolio metrics.
vals <- NULL
for(j in 1:length(AREA)){
  A <- lm(SW_Div~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  B <- lm(SW_Div~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  vals <- rbind(vals,c(AREA[j],
                       coef(summary(A))["Year",c("Estimate","Std. Error")],
                       coef(summary(B))[1,c("Estimate","Std. Error")],
                       var(A$residuals),
                       sd(A$residuals)))
  
}
vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(dat=dat.2,NAME='Shannon Wiener')
```

Simpsons Diversity
```{r, echo=FALSE}
#Plot time series by area  
y.lim <- c(min(dat$simp_Div),max(dat$simp_Div))
for(i in 1:length(AREA)){
  if(AREA[i] !=1){par(new=T)}
  if(AREA[i] >= 3 & AREA[i] <= 5 ){
    plot(simp_Div~Year,data=dat[dat$Area==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
  }
  if(AREA[i] < 3 | AREA[i] > 5){
    plot(simp_Div~Year,data=dat[dat$Area ==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
  }
}
axis(1,at=YEAR,las=2)
axis(2,las=2)
box(bty="o",lwd=2)
title(ylab="Shannon Wiener Index",line=2.5)
#mtext(NAME,adj=0)
abline(v=1989,lty=2,lwd=2)

Calculate least squares trend in each area and portfolio metrics.
vals <- NULL
for(j in 1:length(AREA)){
  A <- lm(simp_Div~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  B <- lm(simp_Div~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  vals <- rbind(vals,c(AREA[j],
                       coef(summary(A))["Year",c("Estimate","Std. Error")],
                       coef(summary(B))[1,c("Estimate","Std. Error")],
                       var(A$residuals),
                       sd(A$residuals)))
  
}
vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(dat=dat.2,NAME='Simpson')
```



OK.  Combine Trends in to one file to make a 4x4 array of plots for potential publication-like Figure.


```{r, echo=FALSE}  
time_plot <- function(dat,NAME){
  AREA <- as.numeric(as.character(unique(dat$Area)))
  YEAR <- sort(unique(dat$Year))
  y.lim <- c(min(dat$Mean.totalDensity),max(dat$Mean.totalDensity))
  for(i in 1:length(AREA)){
    if(AREA[i] !=1){par(new=T)}
    if(AREA[i] >= 3 & AREA[i] <= 5 ){
      plot(Mean.totalDensity~Year,data=dat[dat$Area ==AREA[i],],
            type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
    }
    if(AREA[i] < 3 | AREA[i] > 5){
      plot(Mean.totalDensity~Year,data=dat[dat$Area ==AREA[i],],
            type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
    }
  }
    axis(1,at=YEAR,las=2,hadj=0.6,tcl=-0.25)
    axis(2,las=2,hadj=0.5,tcl=-0.25)
    box(bty="l",lwd=2)
    #title(ylab=expression("Biomass (kg ha"^"-1"*")"),line=2.5)
    mtext(NAME,adj=0.05,line=-1)
    #abline(v=1989,lty=2,lwd=2)
    arrows(x0=1989,x1=1989,y0=y.lim[1],y1=y.lim[1]+c(y.lim[2]-y.lim[1])*0.8,length=0,lty=2,lwd=2)
}

##############################################################################################
# TOTAL BIOMASS
  def.par <- par(no.readonly = TRUE)
  
  PAR <- c(2.5,2,0.5,0.1)
  
  A <- layout(matrix(c(1,2,3,4,1,5,6,7,1,8,9,10,1,11,12,13),4,4,byrow=T),
              widths=c(0.1,1,1,1))
  # Plot shared y axis first
  par(mar=c(0,0,0,0))
  plot(1,1,type="n",xlab="",ylab="",axes=F)
  mtext(expression("Density (kg ha"^"-1"*")"),side=2,line=-2,adj=0.25)
  mtext(expression("Density (kg ha"^"-1"*")"),side=2,line=-2,adj=1)
 
  dat   <- aggregate(meanCPUE[meanCPUE$fish.invert=="fish",c("Mean.totalDensity","vari")],
            by=list(Area=meanCPUE$Area[meanCPUE$fish.invert=="fish"],Year=meanCPUE$year[meanCPUE$fish.invert=="fish"]),sum)
  dat$inv.var <- dat$vari^(-1)
  dat$x.bar <- dat$Mean.totalDensity * dat$inv.var

  par(mar=PAR)
  time_plot(dat,NAME="Total Biomass")

for(i in 1:5){
  par(mar=c(0,0,0,0))
  plot(1,1,type="n",xlab="",ylab="",axes=F)
  mtext(paste("Diversity Metric",i),side=3,line=-5)
}  
  
### DIVERSITY METRIC 


# GUILDS
  GUILD <- c("A","B","P") #unique(meanCPUE$guild)
  for(i in 1:length(GUILD)){
    temp <- meanCPUE[meanCPUE$guild == GUILD[i],]
    dat   <- aggregate(temp[,c("Mean.totalDensity","vari")],by=list(Area=temp$Area,Year=temp$year),sum)
    dat$inv.var <- dat$vari^(-1)
    dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
    par(mar=PAR)
    time_plot(dat,NAME=paste("Guild",GUILD[i]))
  }

# Lifestyle
  DIET <- unique(meanCPUE$diet1)
  for(i in 1:length(DIET)){
    temp <- meanCPUE[meanCPUE$diet1 == DIET[i],]
    dat   <- aggregate(temp[,c("Mean.totalDensity","vari")],by=list(Area=temp$Area,Year=temp$year),sum)
    dat$inv.var <- dat$vari^(-1)
    dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
    time_plot(dat,NAME=paste("Diet",DIET[i]))
  }
```

Repeat for Trend Estimates


