---
title: "Portfolio_Plots_Diversity"
author: "Rachael Blake"
date: "March 21, 2016"
output: html_document
---

```{r, include=FALSE, echo=FALSE, results='hide'}
# Load all necessary packages
library(plyr) ; library(dplyr); library(tidyr) ; library(stringr) ;library(rgdal) ; 
library(sp) ; library(maptools) ; library(raster) ; library(rgeos) ; library(ggplot2) ; 
library(maps) ; library(mapdata) ; library(mapproj) ; library(httr) ; library(extrafont) ; 
library(vegan) ; library(scales)

```

```{r,  include=FALSE, echo=FALSE, results='hide'}
# Load the Species by Area Occurrance data and clean it
URL_SpByArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uWlFSTk03S1pveE0"
SpByArea_Get <- GET(URL_SpByArea)
SpByArea_1 <- content(SpByArea_Get, as='text')
SpByArea <- read.csv(file=textConnection(SpByArea_1), stringsAsFactors=FALSE, header=TRUE)

testing2 <- read.csv("C:/Users/rblake/Documents/NCEAS/GoA Portfolio Effects WG/pfx-groundfish/goaDat.occ.Shallow.AllData.Combined.7.8.9.csv")

meanOcc <- SpByArea %>%
           mutate(Year = ifelse((year.numb=="1"),'1984',
                         ifelse((year.numb=="2"),'1987',
                         ifelse((year.numb=="3"),'1990',
                         ifelse((year.numb=="4"),'1993',       
                         ifelse((year.numb=="5"),'1996',
                         ifelse((year.numb=="6"),'1999',
                         ifelse((year.numb=="7"),'2001',
                         ifelse((year.numb=="8"),'2003',
                         ifelse((year.numb=="9"),'2005',
                         ifelse((year.numb=="10"),'2007',
                         ifelse((year.numb=="11"),'2009',
                         ifelse((year.numb=="12"),'2011',
                         ifelse((year.numb=="13"),'2013',
                         ifelse((year.numb=="14"),'2015',        
                         ""))))))))))))))
                  )
head(meanOcc)

```

```{r,  include=FALSE, echo=FALSE, results='hide'}
# Load the CPUE by Area data and merge in trawl species list
URL_SPCPUEArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-udjdieGNMQ1A5V2M"
SPCPUEArea_Get <- GET(URL_SPCPUEArea)
SPCPUEArea_1 <- content(SPCPUEArea_Get, as='text')
SPCPUEArea <- read.csv(file=textConnection(SPCPUEArea_1), stringsAsFactors=FALSE, header=TRUE)

meanCPUE <- rename(SPCPUEArea, Year=year, Area=area)


```

```{r, include=FALSE, echo=FALSE, results='hide'}
# Read in deeper areas data CPUE
URL_deepCPUE <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-ubjkxNHEzZ0tpN2M"
deepCPUE_Get <- GET(URL_deepCPUE)
deepCPUE_1 <- content(deepCPUE_Get, as='text')
deepCPUE <- read.csv(file=textConnection(deepCPUE_1),stringsAsFactors=FALSE,head=TRUE)

```

```{r, include=FALSE, echo=FALSE, results='hide'}
# Read in deeper areas data OCCURRANCE
URL_deepOCC <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-uV1QtMjZ5bXBka1E"
deepOCC_Get <- GET(URL_deepOCC)
deepOCC_1 <- content(deepOCC_Get, as='text')
deepOCC <- read.csv(file=textConnection(deepOCC_1),stringsAsFactors=FALSE,head=TRUE)

```

```{r, echo=FALSE}
# Make custom theme for plotting
theme_boxplot <- function(base_size = 12){
                 theme_bw(base_size)%+replace%
                 theme(legend.key.size=unit(15,"points"),
                       legend.text=element_text(size=14),
                       legend.key=element_blank(),
                       legend.title=element_blank(),
                       legend.background=element_rect(colour="white", 
                                                      fill="transparent"),
                       plot.margin=unit(c(2,1,2,1), "lines"),  
                       panel.border = element_blank(),
                       panel.margin=unit(0,"lines"),
                       panel.background=element_rect(fill=NA, colour=NA),
                       axis.line.x = element_line(color = "black"),
                       axis.line.y = element_line(color = "black"),
                       axis.ticks.length=unit(1,"mm"),
                       axis.text.x=element_blank(),
                       axis.title.x=element_text(size=17, margin=margin(20,0,0,0)),
                       axis.text.y=element_text(size=15),
                       axis.title.y=element_text(size=17, angle=90,
                                                 margin=margin(0,20,0,0)),
                       panel.grid.major=element_blank(),
                       panel.grid.minor=element_blank(),
                       strip.text.x=element_text(size=14),
                       strip.background=element_rect(colour="black", fill=NA))
}

```

```{r, include=FALSE, echo=FALSE}
# Functions to create plots
weight.var <- function(w){
              return(sum(w)^(-1))
              }

space_plot <- function(df,NAME){
              par(mfrow=c(3,2),mar=c(3,4,1,0.5))
              AT <- df$Area  
              
              plot(1:5,1:5,type="n",axes=F,xlab="",ylab="")
              text(1.5,4,NAME,cex=3,pos=4)
              y.lim <- c(min(df$Trend - df$SE.trend),max(df$Trend + df$SE.trend))
              plot(Trend~Area,data=df,ylim=y.lim,axes=F,pch=21,bg=1)
              abline(h=0,lty=2)
              arrows(x0=df$Area,x1=df$Area,
                     y0=df$Trend + df$SE,y1= df$Trend - df$SE,
                     length=0,lwd=1.5)
              axis(1,at=AT); box(bty="o")
              axis(2,las=2)
              title(xlab="Area",line=2)
              
             # B <- plot(Grand.w.mean~Area,data=df,type="b",axes=F,xlab="",pch=21,bg=1)
          #    axis(1,at=AT); box(bty="o")
          #    axis(2,las=2)
          #    title(xlab="Area",line=2)
              
              plot(SD~Area,data=df,type="b",axes=F,
                   ylim= c(0,max(df$SD)),xlab="",pch=21,bg=1)
              axis(1,at=AT); box(bty="o")
              axis(2,las=2)
              title(xlab="Area",line=2)
      
              plot(CV~Area,data=df,type="b",axes=F,xlab="",pch=21,bg=1)
              axis(1,at=AT); box(bty="o")
              axis(2,las=2)
              title(xlab="Area",line=2)
      
             # E <- plot(Var~Grand.w.mean,data=df,axes=F,xlab="",pch=21,bg=1)
          #    axis(1); box(bty="o")
          #    axis(2,las=2)
          #    title(xlab="W.mean",line=2)
              
             # G <- c(A,C,D)
              
              return()
              
              }



time_plot <- function(df,NAME){
             AREA <- as.numeric(as.character(unique(df$Area)))
             YEAR <- sort(unique(df$Year))
             y.lim <- c(min(df$Mean.totalDensity),max(df$Mean.totalDensity))
             
             for(i in 1:length(AREA)){
                 if(AREA[i] !=1){par(new=T)}
                 if(AREA[i] >= 3 & AREA[i] <= 5 ){
                  plot(Mean.totalDensity~Year,data=df[df$Area ==AREA[i],],
                       type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
                 }
               
             if(AREA[i] < 3 | AREA[i] > 5){
                plot(Mean.totalDensity~Year,data=df[df$Area ==AREA[i],],
                     type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
               }
             }
             
             axis(1,at=YEAR,las=2,hadj=0.6,tcl=-0.25)
             axis(2,las=2,hadj=0.5,tcl=-0.25)
             box(bty="l",lwd=2)
             #title(ylab=expression("Biomass (kg ha"^"-1"*")"),line=2.5)
             mtext(NAME,adj=0.05,line=-1)
             #abline(v=1989,lty=2,lwd=2)
             arrows(x0=1989,x1=1989,y0=y.lim[1],y1=y.lim[1]+
                    c(y.lim[2]-y.lim[1])*0.8,length=0,lty=2,lwd=2)
             }

##########
# Calculate least squares trend in each area and portfolio metrics.

trend_calc <- function(df,y,x, plot_name){
              vals <- NULL

              for(j in 1:length(AREA)){
              A <- lm(y~x,data=df[df$Area == AREA[j] & df$Year > 1989,]) 
              B <- lm(y~1,data=df[df$Area == AREA[j] & df$Year > 1989,]) 
              vals <- rbind(vals,c(AREA[j],
                                   coef(summary(A))["x",c("Estimate","Std. Error")],
                                   coef(summary(B))[1,c("Estimate","Std. Error")],
                                   var(A$residuals),
                                   sd(A$residuals)))
              }

              vals <- data.frame(vals)
              colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean",
                                  "Grand.w.mean.se","Var","SD")
              vals$CV  <- vals$SD / vals$Grand.w.mean 

              df_2 <- vals
              
              C <- space_plot(df=df_2,NAME=plot_name)
              
              return(C)
              }



#trend_calc(df=dat,y="SW_Div",x="Year", plot_name="Shannon Wiener")

#########

Plot_area_time <-function(df,y,ylabel){

                 plot_1 <- ggplot(df, aes(x=Year, y=y, color=Area)) +
                           geom_line(size=1) + theme_bw() + 
                           ylab(ylabel) +
                           geom_vline(aes(xintercept=1989), colour="black",
                                      linetype="dashed") +
                           scale_color_manual(values=line_colors, 
                                              breaks=area_breaks) + 
                           scale_x_continuous(breaks=pretty_breaks(n=15)) + 
                           theme(panel.grid=element_blank(), 
                                 legend.key=element_blank())
     
                 return(plot_1)
                 } 


```


# Need to bootstrap the individual species means per area per year to get a variance
# to include in the diversity calculations - the re-do the diversity calculations
# 
# Look up Kristin France's paper for alpha, beta, gamma diversity and how to plot it up




**Diversity Metrics (biomass data)**

Calculate diversity metrics
```{r, echo=FALSE, include=FALSE}
# Shallower Areas (50-150m)
# First, calculate Shannon and Simpson's using biomass and then calculate Shannon, Simpson and richness using occurrence data (more to be added. e.g. trait and taxonomic divesity)

DivMetrics <- meanCPUE %>%
              filter(Area!="Total") %>%
              group_by(Area, Year) %>% 
              summarise(SW_Div = diversity(Mean.totalDensity, index="shannon"), 
                        Exp_A_Div = exp(SW_Div),  # This is effective number of speices, aka alpha diversity
                        simp_Div = diversity(Mean.totalDensity, index="simpson"), 
                        invsimp = diversity(Mean.totalDensity,index="inv")) %>%
              ungroup() %>%
              print(width = Inf)


dat <- DivMetrics
dat$Area <- as.factor(dat$Area)
AREA <- sort(as.numeric(as.character(unique(dat$Area))))
YEAR <- sort(unique(dat$Year))

```


Here we calculate Shannon diversity, effective # of species, beta diversity, and gamma diversity.
```{r, echo=FALSE}
# Shallower Areas (50-150m)
Div_SpCPUEArea <- meanCPUE %>%
                  filter(Area!="Total") %>%
                  group_by(Area, Year) %>% 
                  mutate(SW_Div = diversity(Mean.totalDensity, index="shannon"),
                         Exp_A_Div = exp(SW_Div)) %>%  # this is alpha diversity
                  ungroup() %>%
                  group_by(Year) %>%
                  mutate(G_Div = diversity(Mean.totalDensity, index="shannon"),
                         Exp_G_Div = exp(G_Div),
                         Exp_B_Div = (Exp_G_Div/Exp_A_Div)) %>%   # calculate beta diversity (gamma = alpha * beta)
                  ungroup() 
              

```

```{r, echo=FALSE}
# Deeper Areas (151-300m)
Div_SpCPUEDeep <- deepCPUE %>%
                  dplyr::select(-X) %>%  
                  mutate_each(funs(as.character), Area) %>%
                  group_by(Area, Year) %>% 
                  mutate(SW_Div = diversity(meanCPUE, index="shannon"),
                         Exp_A_Div = exp(SW_Div)) %>%  # this is alpha diversity
                  ungroup() %>%
                  group_by(Year) %>%
                  mutate(G_Div = diversity(meanCPUE, index="shannon"),
                         Exp_G_Div = exp(G_Div),
                         Exp_B_Div = (Exp_G_Div/Exp_A_Div)) %>%   # calculate beta diversity (gamma = alpha * beta)
                  ungroup() %>%
                  group_by(Area, Year) %>%
                  mutate(Simp_Div = diversity(meanCPUE, index="simpson"), 
                         InvSimp = diversity(meanCPUE,index="inv")) %>%
                  ungroup %>%
                  arrange(Area, Year)


```



Shannon's Index
```{r, echo=FALSE}
#Plot time series by area
y.lim <- c(min(dat$SW_Div),max(dat$SW_Div))

for(i in 1:length(AREA)){
  if(AREA[i] !=1){par(new=T)}
  if(AREA[i] >= 3 & AREA[i] <= 5 ){
    plot(SW_Div~Year,data=dat[dat$Area==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
  }
  if(AREA[i] < 3 | AREA[i] > 5){
    plot(SW_Div~Year,data=dat[dat$Area ==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
  }
}

axis(1,at=YEAR,las=2)
axis(2,las=2)
box(bty="o",lwd=2)
title(ylab="Shannon Wiener Index",line=2.5)
#mtext(NAME,adj=0)
abline(v=1989,lty=2,lwd=2)

# Calculate least squares trend in each area and portfolio metrics.
vals <- NULL

for(j in 1:length(AREA)){
  A <- lm(SW_Div~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  B <- lm(SW_Div~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  vals <- rbind(vals,c(AREA[j],
                       coef(summary(A))["Year",c("Estimate","Std. Error")],
                       coef(summary(B))[1,c("Estimate","Std. Error")],
                       var(A$residuals),
                       sd(A$residuals)))
  
}

vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(df=dat.2,NAME='Shannon Wiener')


#trend_calc(df=dat,y=SW_Div,x=Year, plot_name="Shannon Wiener")

```


Plot of Alpha Diversity "exp(Shannon-Wiener)" ie: effective # of species :   

```{r, echo=FALSE, fig.height=5, fig.width=7}
line_colors <- c("black","black","black","black","red","red","red","black","black",
                 "black","black")
area_breaks <- c("1","2","3","4","5","6","7","8","9","10","11")

Eff_Sp <- Plot_area_time(df=Div_SpCPUEArea, y=Div_SpCPUEArea$Exp_A_Div, 
                         ylabel="Alpha Diversity (exp H')")
Eff_Sp
```

Alpha Diversity deep areas: 
```{r, echo=FALSE, fig.height=5, fig.width=7}
line_colors <- c("red","red","black","black")
area_breaks <- c("12","13","14","15")

Eff_Sp_dp <- Plot_area_time(df=Div_SpCPUEDeep, y=Div_SpCPUEDeep$Exp_A_Div, 
                            ylabel="Alpha Diversity (exp H')")
Eff_Sp_dp


```


Shannon Effective Number exp(H')
```{r, echo=FALSE}
#y.lim <- c(min(dat$Exp_A_Div),max(dat$Exp_A_Div))

#for(i in 1:length(AREA)){
#  if(AREA[i] !=1){par(new=T)}
#  if(AREA[i] >= 3 & AREA[i] <= 5 ){
#    plot(Exp_A_Div~Year,data=dat[dat$Area==AREA[i],],
#         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
#  }
#  if(AREA[i] < 3 | AREA[i] > 5){
#    plot(Exp_A_Div~Year,data=dat[dat$Area ==AREA[i],],
#         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
#  }
#}

#axis(1,at=YEAR,las=2)
#axis(2,las=2)
#box(bty="o",lwd=2)
#title(ylab="exp(H)",line=2.5)
#mtext(NAME,adj=0)
#abline(v=1989,lty=2,lwd=2)

# Calculate least squares trend in each area and portfolio metrics.
vals <- NULL

for(j in 1:length(AREA)){
  A <- lm(Exp_A_Div~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  B <- lm(Exp_A_Div~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  vals <- rbind(vals,c(AREA[j],
                       coef(summary(A))["Year",c("Estimate","Std. Error")],
                       coef(summary(B))[1,c("Estimate","Std. Error")],
                       var(A$residuals),
                       sd(A$residuals)))
  
}

vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(df=dat.2,NAME='exp(H)')

```


Plot Boxplots of Alpha Diversity (exp H'):


```{r, echo=FALSE, fig.height=5, fig.width=7}
# Plotting boxplots of mean of alpha diversity
sp_box <- ggplot(data=Div_SpCPUEArea, aes(x=Area, y=Exp_A_Div)) + 
          geom_boxplot() + theme_boxplot() +  xlab("Area (West <-> East)") +
          ylab("Mean Alpha Diversity (exp H')") +
          xlim("11","10","9","8","7","6","5","4","3","2","1") +
        #  scale_fill_manual(values=barcolor) + ylim(0,0.5) +
          theme(legend.position="none", plot.background=element_blank(),
                axis.text.x=element_text(size=15),
                plot.title=element_text(colour="black", size=15,
                                        hjust=0.04, vjust=0.5, face="bold"))
sp_box
```

```{r, echo=FALSE, fig.height=5, fig.width=7}

sp_box_2 <- ggplot(data=Div_SpCPUEDeep, aes(x=as.factor(Area), y=Exp_A_Div)) + 
            geom_boxplot() + theme_boxplot() +  xlab("Area (West <-> East)") +
            ylab("Mean Alpha Diversity (exp H')") +
            xlim("15","14","13","12") +
          #  scale_fill_manual(values=barcolor) + ylim(0,0.5) +
            theme(legend.position="none", plot.background=element_blank(),
                  axis.text.x=element_text(size=15),
                  plot.title=element_text(colour="black", size=15,
                                          hjust=0.04, vjust=0.5, face="bold"))
sp_box_2


```





Plot beta diversity "(Exp_G_Div/Exp_A_Div)":   

```{r, echo=FALSE, fig.height=5, fig.width=7}
line_colors <- c("black","black","black","black","red","red","red","black","black",
                 "black","black")
area_breaks <- c("1","2","3","4","5","6","7","8","9","10","11")

Bdiv <- Plot_area_time(df=Div_SpCPUEArea, y=Div_SpCPUEArea$Exp_B_Div,
                       ylabel="Beta Diversity")
Bdiv


```

Beta Diversity deep areas: 
```{r, echo=FALSE, fig.height=5, fig.width=7}
line_colors <- c("red","red","black","black")
area_breaks <- c("12","13","14","15")

Bdiv_dp <- Plot_area_time(df=Div_SpCPUEDeep, y=Div_SpCPUEDeep$Exp_B_Div,
                       ylabel="Beta Diversity")
Bdiv_dp

```



Plot gamma diversity:  

```{r, echo=FALSE, fig.height=5, fig.width=7}
Gdiv <- ggplot(Div_SpCPUEArea, aes(Year, Exp_G_Div)) + geom_line(size=1, color="blue") + 
        geom_vline(aes(xintercept=1989), colour="black", linetype="dashed") +
        scale_x_continuous(breaks=pretty_breaks(n=15)) + theme_bw() +
        ylab("Gamma Diversity") +
        theme(panel.grid=element_blank(),
              legend.key=element_blank())
Gdiv

```

Gamma Diversity Deep Areas:
```{r, echo=FALSE, fig.height=5, fig.width=7}
Gdiv <- ggplot(Div_SpCPUEDeep, aes(Year, Exp_G_Div)) + 
        geom_line(size=1, color="blue") + 
        geom_vline(aes(xintercept=1989), colour="black", linetype="dashed") +
        scale_x_continuous(breaks=pretty_breaks(n=15)) + theme_bw() +
        ylab("Gamma Diversity") +
        theme(panel.grid=element_blank(),
              legend.key=element_blank())
Gdiv

```



Simpson's Diversity
```{r, echo=FALSE}
### Simpsons Diversity
#Plot time series by area  
y.lim <- c(min(dat$simp_Div),max(dat$simp_Div))
for(i in 1:length(AREA)){
  if(AREA[i] !=1){par(new=T)}
  if(AREA[i] >= 3 & AREA[i] <= 5 ){
    plot(simp_Div~Year,data=dat[dat$Area==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
  }
  if(AREA[i] < 3 | AREA[i] > 5){
    plot(simp_Div~Year,data=dat[dat$Area ==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
  }
}
axis(1,at=YEAR,las=2)
axis(2,las=2)
box(bty="o",lwd=2)
title(ylab="Simpson",line=2.5)
#mtext(NAME,adj=0)
abline(v=1989,lty=2,lwd=2)

#Calculate least squares trend in each area and portfolio metrics.
vals <- NULL

for(j in 1:length(AREA)){
    A <- lm(invsimp~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
    B <- lm(invsimp~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
    vals <- rbind(vals,c(AREA[j],
                         coef(summary(A))["Year",c("Estimate","Std. Error")],
                         coef(summary(B))[1,c("Estimate","Std. Error")],
                         var(A$residuals),
                         sd(A$residuals)))
  
}

vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(df=dat.2,NAME='Simpson Index')

```


Inverse Simpsons
```{r, echo=FALSE}
### Inverse Simpsons Diversity
#Plot time series by area  
y.lim <- c(min(dat$invsimp),max(dat$invsimp))

for(i in 1:length(AREA)){
  if(AREA[i] !=1){par(new=T)}
  if(AREA[i] >= 3 & AREA[i] <= 5 ){
    plot(invsimp~Year,data=dat[dat$Area==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
  }
  if(AREA[i] < 3 | AREA[i] > 5){
    plot(invsimp~Year,data=dat[dat$Area ==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
  }
}

axis(1,at=YEAR,las=2)
axis(2,las=2)
box(bty="o",lwd=2)
title(ylab="Inverse Simpson",line=2.5)
#mtext(NAME,adj=0)
abline(v=1989,lty=2,lwd=2)

#Calculate least squares trend in each area and portfolio metrics.
vals <- NULL
for(j in 1:length(AREA)){
  A <- lm(invsimp~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  B <- lm(invsimp~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  vals <- rbind(vals,c(AREA[j],
                       coef(summary(A))["Year",c("Estimate","Std. Error")],
                       coef(summary(B))[1,c("Estimate","Std. Error")],
                       var(A$residuals),
                       sd(A$residuals)))
  
}

vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(df=dat.2,NAME='Inv Simpson')

```




**Diversity Metrics (occurrence data)**

Calculate diversity metrics
```{r, echo=FALSE, include=FALSE}  
DivMetrics.Occ <- meanOcc %>%
                  rename(Area=area) %>%
                  filter(Area!="Total") %>%
                  group_by(Area, Year) %>% 
                  summarise(SW_Div = diversity(Mean.avgPresence, index="shannon"), 
                            Eff_Num_Sp = exp(SW_Div),
                            simp_Div = diversity(Mean.avgPresence, index="simpson"), 
                            invsimp = diversity(Mean.avgPresence,index="inv"),
                            SpRich = sum(Mean.avgPresence)) %>%
                  ungroup() %>%
                  print(width = Inf)

dat_occ <- DivMetrics.Occ
dat_occ$Year <- as.numeric(dat_occ$Year)
dat_occ$Area <- as.factor(dat_occ$Area)
AREA <- sort(as.numeric(as.character(unique(dat_occ$Area))))
YEAR <- sort(unique(dat_occ$Year))
```

Here we calculate Shannon diversity, effective # of species, beta diversity, and gamma diversity.
```{r, echo=FALSE}

Div_SpOccArea <- meanOcc %>%
                 mutate_each(funs(as.numeric), Year) %>%
                 rename(Area=area) %>%
                 filter(Area!="Total") %>%
                 group_by(Area, Year) %>% 
                 mutate(SW_Div = diversity(Mean.avgPresence, index="shannon"),
                        Exp_A_Div = exp(SW_Div), 
                        Sp_Rich = sum(Mean.avgPresence)) %>%  # this is alpha diversity
                 ungroup() %>%
                 group_by(Year) %>%
                 mutate(G_Div = diversity(Mean.avgPresence, index="shannon"),
                        Exp_G_Div = exp(G_Div),
                        Exp_B_Div = (Exp_G_Div/Exp_A_Div)) %>%   # calculate beta diversity (gamma = alpha * beta)
                 ungroup() 
              

```

Species Richness by Area.
```{r, echo=FALSE, fig.height=5, fig.width=7}
# Shallow Species Richness
line_colors <- c("black","black","black","black","red","red","red","black","black",
                 "black","black")
area_breaks <- c("1","2","3","4","5","6","7","8","9","10","11")

SpRdiv <- Plot_area_time(df=Div_SpOccArea, y=Div_SpOccArea$Sp_Rich,
                         ylabel="Species Richness")
SpRdiv


```

```{r, echo=FALSE, fig.height=5, fig.width=7}
# Deep Species Richness


```


Species richness
```{r, echo=FALSE}  
#Species richness
#Plot time series by area
#y.lim <- c(min(dat_occ$SpRich),max(dat_occ$SpRich))

#for(i in 1:length(AREA)){
#  if(AREA[i] !=1){par(new=T)}
#  if(AREA[i] >= 3 & AREA[i] <= 5 ){
#    plot(SpRich~Year,data=dat_occ[dat_occ$Area==AREA[i],],
#         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
#  }
#  if(AREA[i] < 3 | AREA[i] > 5){
#    plot(SpRich~Year,data=dat_occ[dat_occ$Area ==AREA[i],],
#         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
#  }
#}

#axis(1,at=YEAR,las=2)
#axis(2,las=2)
#box(bty="o",lwd=2)
#title(ylab="Species Richness",line=2.5)
#mtext(NAME,adj=0)
#abline(v=1989,lty=2,lwd=2)

# Calculate least squares trend in each area and portfolio metrics.
vals <- NULL

for(j in 1:length(AREA)){
  A <- lm(SpRich~Year,data=dat_occ[dat_occ$Area == AREA[j] & dat_occ$Year > 1989,]) #,weights = inv.var )
  B <- lm(SpRich~1,data=dat_occ[dat_occ$Area == AREA[j] & dat_occ$Year > 1989,]) #,weights = inv.var )
  vals <- rbind(vals,c(AREA[j],
                       coef(summary(A))["Year",c("Estimate","Std. Error")],
                       coef(summary(B))[1,c("Estimate","Std. Error")],
                       var(A$residuals),
                       sd(A$residuals)
                       )
                  )
}

vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(df=dat.2,NAME="Species Richness")

```









```{r, echo=FALSE, eval=FALSE}  
##############################################################################################
# TOTAL BIOMASS
  def.par <- par(no.readonly = TRUE)
  
  PAR <- c(2.5,2,0.5,0.1)
  
  A <- layout(matrix(c(1,2,3,4,1,5,6,7,1,8,9,10,1,11,12,13),4,4,byrow=T),
              widths=c(0.1,1,1,1))
  
  # Plot shared y axis first
  par(mar=c(0,0,0,0))
  plot(1,1,type="n",xlab="",ylab="",axes=F)
  mtext(expression("Density (kg ha"^"-1"*")"),side=2,line=-2,adj=0.25)
  mtext(expression("Density (kg ha"^"-1"*")"),side=2,line=-2,adj=1)
 
  dat   <- aggregate(meanCPUE[meanCPUE$fish.invert=="fish",c("Mean.totalDensity","vari")],
            by=list(Area=meanCPUE$Area[meanCPUE$fish.invert=="fish"],
                    Year=meanCPUE$year[meanCPUE$fish.invert=="fish"]),sum)
  dat$inv.var <- dat$vari^(-1)
  dat$x.bar <- dat$Mean.totalDensity * dat$inv.var

  par(mar=PAR)
  time_plot(dat,NAME="Total Biomass")

for(i in 1:5){
  par(mar=c(0,0,0,0))
  plot(1,1,type="n",xlab="",ylab="",axes=F)
  mtext(paste("Diversity Metric",i),side=3,line=-5)
}  

```    
  
  
```{r, echo=FALSE, eval=FALSE}  
### DIVERSITY METRIC 
# GUILDS
  GUILD <- c("A","B","P") #unique(meanCPUE$guild)
  for(i in 1:length(GUILD)){
    temp <- meanCPUE[meanCPUE$guild == GUILD[i],]
    dat   <- aggregate(temp[,c("Mean.totalDensity","vari")],by=list(Area=temp$Area,Year=temp$year),sum)
    dat$inv.var <- dat$vari^(-1)
    dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
    par(mar=PAR)
    time_plot(dat,NAME=paste("Guild",GUILD[i]))
  }
```

```{r, echo=FALSE, eval=FALSE}
# Lifestyle
  DIET <- unique(meanCPUE$diet1)
  for(i in 1:length(DIET)){
    temp <- meanCPUE[meanCPUE$diet1 == DIET[i],]
    dat   <- aggregate(temp[,c("Mean.totalDensity","vari")],by=list(Area=temp$Area,Year=temp$year),sum)
    dat$inv.var <- dat$vari^(-1)
    dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
    time_plot(dat,NAME=paste("Diet",DIET[i]))
  }
```

Repeat for Trend Estimates



