---
title: "Portfolio_Plots_Diversity"
author: "Rachael Blake"
date: "March 21, 2016"
output: html_document
---

```{r, include=FALSE, echo=FALSE, results='hide'}
library(plyr) ; library(dplyr); library(tidyr) ; library(stringr) ;library(rgdal) ; 
library(sp) ; library(maptools) ; library(raster) ; library(rgeos) ; library(ggplot2) ; 
library(maps) ; library(mapdata) ; library(mapproj) ; library(httr) ; library(extrafont) ; 
library(vegan)

```

```{r,  include=FALSE, echo=FALSE, results='hide'}
URL_SpByArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-udlVNME9rQXEwZ1k"
SpByArea_Get <- GET(URL_SpByArea)
SpByArea_1 <- content(SpByArea_Get, as='text')
SpByArea <- read.csv(file=textConnection(SpByArea_1),stringsAsFactors=FALSE,head=TRUE)


meanOcc <- SpByArea %>%
            mutate(Year = ifelse((year.numb=="1"),'1984',
                          ifelse((year.numb=="2"),'1987',
                          ifelse((year.numb=="3"),'1990',
                          ifelse((year.numb=="4"),'1993',       
                          ifelse((year.numb=="5"),'1996',
                          ifelse((year.numb=="6"),'1999',
                          ifelse((year.numb=="7"),'2001',
                          ifelse((year.numb=="8"),'2003',
                          ifelse((year.numb=="9"),'2005',
                          ifelse((year.numb=="10"),'2007',
                          ifelse((year.numb=="11"),'2009',
                          ifelse((year.numb=="12"),'2011',
                          ifelse((year.numb=="13"),'2013',
                          ifelse((year.numb=="14"),'2015',        
                          ""))))))))))))))
                   )
head(meanOcc)

```

```{r,  include=FALSE, echo=FALSE, results='hide'}
URL_SPCPUEArea <- "https://drive.google.com/uc?export=download&id=0By1iaulIAI-udm1FT2trQUh5N1k"
SPCPUEArea_Get <- GET(URL_SPCPUEArea)
SPCPUEArea_1 <- content(SPCPUEArea_Get, as='text')
SPCPUEArea <- read.csv(file=textConnection(SPCPUEArea_1),stringsAsFactors=FALSE,head=TRUE)

meanCPUE <- rename(SPCPUEArea, Year=year, Area=area)

```


```{r, echo=FALSE}
# Functions to create plots
weight.var <- function(w){
  return(sum(w)^(-1))
}
space_plot <- function(dat,NAME){
    par(mfrow=c(3,2),mar=c(3,4,1,0.5))
    AT <- dat$Area  
    plot(1:5,1:5,type="n",axes=F,xlab="",ylab="")
    text(1.5,4,NAME,cex=3,pos=4)
    y.lim <- c(min(dat$Trend - dat$SE.trend),max(dat$Trend + dat$SE.trend))
    plot(Trend~Area,data=dat,ylim=y.lim,axes=F,pch=21,bg=1)
    abline(h=0,lty=2)
    arrows(x0=dat$Area,x1=dat$Area,
          y0=dat$Trend + dat$SE,y1= dat$Trend - dat$SE,
          length=0,lwd=1.5)
      axis(1,at=AT); box(bty="o")
      axis(2,las=2)
      title(xlab="Area",line=2)
    plot(Grand.w.mean~Area,data=dat,type="b",axes=F,xlab="",pch=21,bg=1)
      axis(1,at=AT); box(bty="o")
      axis(2,las=2)
      title(xlab="Area",line=2)
    plot(SD~Area,data=dat,type="b",axes=F,ylim= c(0,max(dat$SD)),xlab="",pch=21,bg=1)
      axis(1,at=AT); box(bty="o")
      axis(2,las=2)
      title(xlab="Area",line=2)
    plot(CV~Area,data=dat,type="b",axes=F,xlab="",pch=21,bg=1)
      axis(1,at=AT); box(bty="o")
      axis(2,las=2)
      title(xlab="Area",line=2)
    plot(Var~Grand.w.mean,data=dat,axes=F,xlab="",pch=21,bg=1)
      axis(1); box(bty="o")
      axis(2,las=2)
      title(xlab="W.mean",line=2)
}
```


**Diversity Metrics (biomass data)**

Calculate diversity metrics
```{r, echo=FALSE}
# First, calculate Shannon and Simpson's using biomass and then calculate Shannon, Simpson and richness using occurrence data (more to be added. e.g. trait and taxonomic divesity)


# Metrics using Biomass data

DivMetrics <- meanCPUE %>%
              filter(Area!="Total") %>%
              group_by(Area, Year) %>% 
              summarise(SW_Div = diversity(Mean.totalDensity, index="shannon"), 
                        Exp_A_Div = exp(SW_Div),  # This is effective number of speices, aka alpha diversity
                        simp_Div = diversity(Mean.totalDensity, index="simpson"), 
                        invsimp = diversity(Mean.totalDensity,index="inv")) %>%
              ungroup() %>%
              print(width = Inf)


dat <- DivMetrics
AREA <- as.numeric(as.character(unique(dat$Area)))
YEAR <- sort(unique(dat$Year))
```


Here we calculate Shannon diversity, effective # of species, beta diversity, and gamma diversity.
```{r, echo=FALSE}
Div_SpCPUEArea <- meanCPUE %>%
                  filter(Area!="Total") %>%
                  group_by(area, Year) %>% 
                  mutate(SW_Div = diversity(Mean.totalDensity, index="shannon"),
                         Exp_A_Div = exp(SW_Div)) %>%  # this is alpha diversity
                  ungroup() %>%
                  group_by(Year) %>%
                  mutate(G_Div = diversity(Mean.totalDensity, index="shannon"),
                         Exp_G_Div = exp(G_Div),
                         Exp_B_Div = (Exp_G_Div/Exp_A_Div)) %>%   # calculate beta diversity (gamma = alpha * beta)
                  ungroup() 
              

```


Shannon's Index
```{r, echo=FALSE}
#Plot time series by area
y.lim <- c(min(dat$Eff_Num_Sp),max(dat$Eff_Num_Sp))
for(i in 1:length(AREA)){
  if(AREA[i] !=1){par(new=T)}
  if(AREA[i] >= 3 & AREA[i] <= 5 ){
    plot(Eff_Num_Sp~Year,data=dat[dat$Area==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
  }
  if(AREA[i] < 3 | AREA[i] > 5){
    plot(Eff_Num_Sp~Year,data=dat[dat$Area ==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
  }
}
axis(1,at=YEAR,las=2)
axis(2,las=2)
box(bty="o",lwd=2)
title(ylab="Shannon's Diversity",line=2.5)
#mtext(NAME,adj=0)
abline(v=1989,lty=2,lwd=2)

#Calculate least squares trend in each area and portfolio metrics
vals <- NULL
for(j in 1:length(AREA)){
  A <- lm(Eff_Num_Sp~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  B <- lm(Eff_Num_Sp~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  vals <- rbind(vals,c(AREA[j],
                       coef(summary(A))["Year",c("Estimate","Std. Error")],
                       coef(summary(B))[1,c("Estimate","Std. Error")],
                       var(A$residuals),
                       sd(A$residuals)))
  
}
vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(dat=dat.2,NAME='Shannon's Diversity')
```

Simpson's Diversity
```{r, echo=FALSE}
#Plot time series by area  
y.lim <- c(min(dat$invsimp),max(dat$invsimp))
for(i in 1:length(AREA)){
  if(AREA[i] !=1){par(new=T)}
  if(AREA[i] >= 3 & AREA[i] <= 5 ){
    plot(invsimp~Year,data=dat[dat$Area==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
  }
  if(AREA[i] < 3 | AREA[i] > 5){
    plot(invsimp~Year,data=dat[dat$Area ==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
  }
}
axis(1,at=YEAR,las=2)
axis(2,las=2)
box(bty="o",lwd=2)
title(ylab="Simpson's Index",line=2.5)
#mtext(NAME,adj=0)
abline(v=1989,lty=2,lwd=2)

#Calculate least squares trend in each area and portfolio metrics.
vals <- NULL
for(j in 1:length(AREA)){
  A <- lm(invsimp~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  B <- lm(invsimp~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  vals <- rbind(vals,c(AREA[j],
                       coef(summary(A))["Year",c("Estimate","Std. Error")],
                       coef(summary(B))[1,c("Estimate","Std. Error")],
                       var(A$residuals),
                       sd(A$residuals)))
  
}
vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(dat=dat.2,NAME='Simpson's Index')
```

**Diversity Metrics (occurrence data)**

Calculate diversity metrics
```{r, echo=FALSE}  
DivMetrics.Occ <- meanOcc %>%
  group_by(Area, Year) %>% 
  summarise(SW_Div = diversity(Mean.avgPresence, index="shannon"), Eff_Num_Sp = exp(SW_Div),
            simp_Div = diversity(Mean.avgPresence, index="simpson"), 
            invsimp = diversity(Mean.avgPresence,index="inv"),SpRich = sum(Mean.avgPresence)) %>%
  ungroup() %>%
  print(width = Inf)

dat<-DivMetrics.Occ
AREA <- as.numeric(as.character(unique(dat$Area)))
YEAR <- sort(unique(dat$Year))
```

Shannon Wiener
```{r, echo=FALSE}  
#Plot time series by area
y.lim <- c(min(dat$Eff_Num_Sp),max(dat$Eff_Num_Sp))
for(i in 1:length(AREA)){
  if(AREA[i] !=1){par(new=T)}
  if(AREA[i] >= 3 & AREA[i] <= 5 ){
    plot(Eff_Num_Sp~Year,data=dat[dat$Area==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
  }
  if(AREA[i] < 3 | AREA[i] > 5){
    plot(Eff_Num_Sp~Year,data=dat[dat$Area ==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
  }
}
axis(1,at=YEAR,las=2)
axis(2,las=2)
box(bty="o",lwd=2)
title(ylab="Shannon's Diversity",line=2.5)
#mtext(NAME,adj=0)
abline(v=1989,lty=2,lwd=2)

# Calculate least squares trend in each area and portfolio metrics.
vals <- NULL
for(j in 1:length(AREA)){
  A <- lm(Eff_Num_Sp~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  B <- lm(Eff_Num_Sp~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  vals <- rbind(vals,c(AREA[j],
                       coef(summary(A))["Year",c("Estimate","Std. Error")],
                       coef(summary(B))[1,c("Estimate","Std. Error")],
                       var(A$residuals),
                       sd(A$residuals)))
  
}
vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(dat=dat.2,NAME="Shannon's Diversity")
```

Simpson's Diversity
```{r, echo=FALSE}  
#Plot time series by area
y.lim <- c(min(dat$invsimp),max(dat$invsimp))
for(i in 1:length(AREA)){
  if(AREA[i] !=1){par(new=T)}
  if(AREA[i] >= 3 & AREA[i] <= 5 ){
    plot(invsimp~Year,data=dat[dat$Area==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
  }
  if(AREA[i] < 3 | AREA[i] > 5){
    plot(invsimp~Year,data=dat[dat$Area ==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
  }
}
axis(1,at=YEAR,las=2)
axis(2,las=2)
box(bty="o",lwd=2)
title(ylab="Simpson's Index",line=2.5)
#mtext(NAME,adj=0)
abline(v=1989,lty=2,lwd=2)

# Calculate least squares trend in each area and portfolio metrics.
vals <- NULL
for(j in 1:length(AREA)){
  A <- lm(invsimp~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  B <- lm(invsimp~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  vals <- rbind(vals,c(AREA[j],
                       coef(summary(A))["Year",c("Estimate","Std. Error")],
                       coef(summary(B))[1,c("Estimate","Std. Error")],
                       var(A$residuals),
                       sd(A$residuals)))
  
}
vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(dat=dat.2,NAME="Simpson's Index")
``` 

Species richness
```{r, echo=FALSE}  
#Plot time series by area
y.lim <- c(min(dat$SpRich),max(dat$SpRich))
for(i in 1:length(AREA)){
  if(AREA[i] !=1){par(new=T)}
  if(AREA[i] >= 3 & AREA[i] <= 5 ){
    plot(SpRich~Year,data=dat[dat$Area==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
  }
  if(AREA[i] < 3 | AREA[i] > 5){
    plot(SpRich~Year,data=dat[dat$Area ==AREA[i],],
         type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
  }
}
axis(1,at=YEAR,las=2)
axis(2,las=2)
box(bty="o",lwd=2)
title(ylab="Species Richness",line=2.5)
#mtext(NAME,adj=0)
abline(v=1989,lty=2,lwd=2)

# Calculate least squares trend in each area and portfolio metrics.
vals <- NULL
for(j in 1:length(AREA)){
  A <- lm(SpRich~Year,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  B <- lm(SpRich~1,data=dat[dat$Area == AREA[j] & dat$Year > 1989,]) #,weights = inv.var )
  vals <- rbind(vals,c(AREA[j],
                       coef(summary(A))["Year",c("Estimate","Std. Error")],
                       coef(summary(B))[1,c("Estimate","Std. Error")],
                       var(A$residuals),
                       sd(A$residuals)))
  
}
vals <- data.frame(vals)
colnames(vals) <- c("Area","Trend","SE.trend","Grand.w.mean","Grand.w.mean.se","Var","SD")
vals$CV  <- vals$SD / vals$Grand.w.mean 

dat.2 <- vals

space_plot(dat=dat.2,NAME="Species Richness")

```


OK.  Combine Trends in to one file to make a 4x4 array of plots for potential publication-like Figure.
```{r, echo=FALSE}  
time_plot <- function(dat,NAME){
  AREA <- as.numeric(as.character(unique(dat$Area)))
  YEAR <- sort(unique(dat$Year))
  y.lim <- c(min(dat$Mean.totalDensity),max(dat$Mean.totalDensity))
  for(i in 1:length(AREA)){
    if(AREA[i] !=1){par(new=T)}
    if(AREA[i] >= 3 & AREA[i] <= 5 ){
      plot(Mean.totalDensity~Year,data=dat[dat$Area ==AREA[i],],
            type="l",axes=F,xlab="",ylim=y.lim,col=2,lwd=2,ylab="")
    }
    if(AREA[i] < 3 | AREA[i] > 5){
      plot(Mean.totalDensity~Year,data=dat[dat$Area ==AREA[i],],
            type="l",axes=F,xlab="",ylim=y.lim,col=1,lwd=1.5,ylab="")
    }
  }
    axis(1,at=YEAR,las=2,hadj=0.6,tcl=-0.25)
    axis(2,las=2,hadj=0.5,tcl=-0.25)
    box(bty="l",lwd=2)
    #title(ylab=expression("Biomass (kg ha"^"-1"*")"),line=2.5)
    mtext(NAME,adj=0.05,line=-1)
    #abline(v=1989,lty=2,lwd=2)
    arrows(x0=1989,x1=1989,y0=y.lim[1],y1=y.lim[1]+c(y.lim[2]-y.lim[1])*0.8,length=0,lty=2,lwd=2)
}

##############################################################################################
# TOTAL BIOMASS
  def.par <- par(no.readonly = TRUE)
  
  PAR <- c(2.5,2,0.5,0.1)
  
  A <- layout(matrix(c(1,2,3,4,1,5,6,7,1,8,9,10,1,11,12,13),4,4,byrow=T),
              widths=c(0.1,1,1,1))
  # Plot shared y axis first
  par(mar=c(0,0,0,0))
  plot(1,1,type="n",xlab="",ylab="",axes=F)
  mtext(expression("Density (kg ha"^"-1"*")"),side=2,line=-2,adj=0.25)
  mtext(expression("Density (kg ha"^"-1"*")"),side=2,line=-2,adj=1)
 
  dat   <- aggregate(meanCPUE[meanCPUE$fish.invert=="fish",c("Mean.totalDensity","vari")],
            by=list(Area=meanCPUE$Area[meanCPUE$fish.invert=="fish"],Year=meanCPUE$year[meanCPUE$fish.invert=="fish"]),sum)
  dat$inv.var <- dat$vari^(-1)
  dat$x.bar <- dat$Mean.totalDensity * dat$inv.var

  par(mar=PAR)
  time_plot(dat,NAME="Total Biomass")

for(i in 1:5){
  par(mar=c(0,0,0,0))
  plot(1,1,type="n",xlab="",ylab="",axes=F)
  mtext(paste("Diversity Metric",i),side=3,line=-5)
}  
  
### DIVERSITY METRIC 


# GUILDS
  GUILD <- c("A","B","P") #unique(meanCPUE$guild)
  for(i in 1:length(GUILD)){
    temp <- meanCPUE[meanCPUE$guild == GUILD[i],]
    dat   <- aggregate(temp[,c("Mean.totalDensity","vari")],by=list(Area=temp$Area,Year=temp$year),sum)
    dat$inv.var <- dat$vari^(-1)
    dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
    par(mar=PAR)
    time_plot(dat,NAME=paste("Guild",GUILD[i]))
  }

# Lifestyle
  DIET <- unique(meanCPUE$diet1)
  for(i in 1:length(DIET)){
    temp <- meanCPUE[meanCPUE$diet1 == DIET[i],]
    dat   <- aggregate(temp[,c("Mean.totalDensity","vari")],by=list(Area=temp$Area,Year=temp$year),sum)
    dat$inv.var <- dat$vari^(-1)
    dat$x.bar <- dat$Mean.totalDensity * dat$inv.var
  
    time_plot(dat,NAME=paste("Diet",DIET[i]))
  }
```

Repeat for Trend Estimates



