---
title: "Univariate plotting & summary metrics of spatial model output"
author: "Ben Williams"
date: "October 26, 2015"
output: html_document
---
This is the initial examination using Ole's model output file from 2015-10-29.

The *vplot* function will generate a figure for each year and area. Input a dataframe containing the species of interest, subsets of the "All_sp_index_meanCPUEByArea.csv"" file. The location of this may change when the model is updated.

A seperate file can be used to house a data.frame of species names and trophic guilds etc. This file can be used to inform vplot of the groups to plot (example code below). Further this will allow for relatively painless exmainations of guild variance vs the whole, etc.

```{r load, message=FALSE, echo=FALSE}
library(extrafont)
#font_import() only do this one time - it takes a while
loadfonts(device="win")
windowsFonts(Times=windowsFont("TT Times New Roman"))

library(ggplot2)
theme_set(theme_bw(base_size=12,base_family='Times New Roman')+ 
  theme(panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()))

library(maps)
library(mapdata)
library(mapproj)
```

```{r, cache=FALSE, echo=FALSE}
meanCPUE <- read.csv("All_sp_index_meanCPUEByArea.csv") # load data
meanCPUE <- subset(meanCPUE,area!='Total') #remove total field
meanCPUE <- droplevels(meanCPUE) 
meanCPUE$area <- as.numeric(levels(meanCPUE$area))[meanCPUE$area]#use numeric/factor conventions
meanCPUE$Area <- factor(meanCPUE$area)
meanCPUE$vari <- meanCPUE$SD.totalDensity^2 #calculate variance

discrete_areas <- read.csv("./goaTrawl/Output Data/goa_discrete_areas_for_comparison(50_to_150m).csv")
#head(discrete_areas)

trawl_species <- read.csv("./goaTrawl/Output Data/trawl_species_control_file.csv")
```

```{r map, cache=FALSE, echo=FALSE}
ak<-map_data('worldHires','USA:Alaska')
akmap<-ggplot()+geom_polygon(data=ak,aes(long,lat,group=group),fill=8,color="black")+
    theme(panel.background=element_rect(fill='aliceblue'))+
    xlab(expression(paste(Longitude^o,~'W')))+
    ylab(expression(paste(Latitude^o,~'N')))+
    coord_map(xlim = c(-160, -143),ylim = c(54, 61))
```

Map of area locations

```{r, echo=FALSE, cache=FALSE}
discrete_areas$area <- discrete_areas$Area
discrete_areas$Area <- factor(discrete_areas$area)
akmap+geom_point(data=discrete_areas,aes(LONGITUDE,LATITUDE,color=Area))
```

Function for generating figures

```{r vplot,echo=TRUE}
vplot <- function(x, title=""){
a <- aggregate(vari~area+year,data=x,FUN=sum)
b <- aggregate(Mean.totalDensity~area+year,data=x,FUN=sum)
c <- merge(a, b, by=c("area", "year"))
c$se <- sqrt(c$vari)
c$uci <- c$se*2 + c$Mean.totalDensity
c$lci <- c$Mean.totalDensity - c$se*2
c$Area <- factor(c$area)

ggplot(c, aes(year, Mean.totalDensity, group=Area, color=Area))+geom_line()+
  geom_ribbon(aes(ymin=lci, ymax=uci, fill=Area), alpha=.2, linetype=0)+
  ylim(min(c$lci), max(c$uci))+ scale_x_continuous(limits=c(1984, 2012), breaks=seq(1984,2012,2))+ggtitle(title)
}
```

Subset groups based upon "trawl_species_control_file.csv"
```{r}
sp <- subset(trawl_species, pelagic.benthic!='pelagic')
sp <- droplevels(sp)
sp.names <- as.list(levels(sp$database.name))
```

Define the group to plot based upon the species names pulled from the database
```{r}
group <- meanCPUE[meanCPUE$Species %in% sp.names,]
```

Plot the group

```{r}
vplot(group,title="Demersal sp.")
```

Examine by guild
```{r guild}
sp <- subset(trawl_species, guild!='B')
sp <- droplevels(sp)
sp.names <- as.list(levels(sp$database.name))

group <- meanCPUE[meanCPUE$Species %in% sp.names,]

vplot(group, title="Guild B")
```


Plot of the sum and variance of mean total density by year and area
```{r}
vplot(meanCPUE, title="All species")
```

Area 8 jumps out as higher than the other areas


Examine all species minus Walleye Pollock, Arrowtooth Flounder and Pacific Halibut

```{r, echo=TRUE}
#subset by area
smean <- subset(meanCPUE, Species!="Theragrachalcogramma" & Species!="Atheresthesstomias"& Species!="Hippoglossoideselassodon")
smean <- droplevels(smean)

vplot(smean)
```


Walleye Pollock, Arrowtooth Flounder and Pacific Halibut

```{r, echo=FALSE}
polarrhal <- subset(meanCPUE, Species=="Theragrachalcogramma" | Species== "Atheresthesstomias" | Species!="Hippoglossoideselassodon" )
polarrhal <- droplevels(polarrhal)

vplot(polarrhal)
```

Cod and Spiny Dogfish

```{r, echo=FALSE}
codsha <- subset(meanCPUE, Species=="Gadusmacrocephalus"| Species=="Squalusacanthias")

vplot(codsha)
```

Capelin, Herring, and Eulachon

```{r, echo=FALSE}
capherreu <- subset(meanCPUE,Species== "Mallotusvillosus"|Species== "Clupeapallasi" | Species=="Thaleichthyspacificus")

vplot(capherreu)
```

Rock Sole, Yellowfin Sole, Bigmouth Sculpin, Ling Cod

```{r, echo=FALSE}
#pull out species "groups"
rocyelbig <- subset(meanCPUE, Species=="Lepidopsettasp"|Species== "Limandaaspera"|Species== "Hemitripterusbolini" | Species == "Ophiodonelongatus" |Species =="Dasycottussetiger" | Species=="Hemilepidotus Jordani")

vplot(rocyelbig)
```

Bairdi Crab

```{r, echo=FALSE}
bai <- subset(meanCPUE, Species=="Chionoecetesbairdi")

vplot(bai)
```
